---
title: "Data Preparation for National FADN and LPIS"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Data Preparation for National FADN and LPIS}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
bibliography: references.bib
link-citations: true
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)

library(FADN2Footprint)

library(dplyr)
library(tidyr)
library(readxl)
library(gt)
```

## Overview

National FADN databases differ from the EU FADN in three ways:

1. **Table structure** — National FADNs use **long product tables** (one row per
   farm × product) while the EU FADN is a **single wide table** (one row per farm).
2. **Variable naming** — Different suffixes encode the same concept (e.g., `SUPER3`
   in RICA = `TA` in FADN = total area).
3. **Category codes** — Products are identified by national codes (e.g., RICA code
   `111` = FADN code `CWHTC`).

The conversion is driven by **two user-filled tables**:

| Table | What it maps | Example |
|---|---|---|
| **Variable pattern** | Suffix ↔ suffix within a product group | `SUPER3` → `TA` (total area, CROPS) |
| **Category pattern** | National product code → FADN product code | `111` → `CWHTC` (common wheat) |

Together they generate the full variable mapping: national long-table column
`SUPER3` for code `111` becomes wide-table column `CWHTC_TA`.

---

## Understanding the two conversion tables

### Variable pattern table

Each row maps one **variable suffix** in the national FADN to one **variable
suffix** in the EU FADN, within a product group. Some FADN suffixes have no
national equivalent (and vice versa) — these rows are left with the
corresponding cell empty.

Here is the structure the user must fill:

| Column | Description | Example |
|---|---|---|
| `pattern_FADN` | EU FADN variable suffix | `TA` |
| `description_FADN` | Description of the EU FADN variable | Total area under production |
| `Comment` | Unit or clarification | in ha |
| `Group` | Product group: `CROPS`, `LIVESTOCK`, or `ANIMAL PRODUCTS AND SERVICES` | `CROPS` |
| `pattern_national` | National (RICA) suffix in the long table | `SUPER3` |
| `description_national` | Description of the national variable | Surface |

**Key observations from the RICA example:**

- Some FADN suffixes have **no RICA equivalent** (e.g., `A` = AA for crops,
  `ENA` = energy crop area, `GM` = GMO area, `ENTA` = energy crop total area,
  `IRA` = irrigated AA, many livestock suffixes like `SSN`, `SSV`, `SRN`, `SRV`,
  `NO`, `FCN`, `FCV`, `FUN`, `FUV`, `PI`, `ALU`). These are left with an empty
  `pattern_national` cell.
- Some RICA suffixes have **no FADN equivalent** — these can be included with an
  empty `pattern_FADN` cell if the user wants to preserve them.
- The `Group` column determines which raw table and code column to use:
  - `CROPS` → table `veg20`, code column `CODE3`
  - `LIVESTOCK` → table `ani20`, code column `CODE6`
  - `ANIMAL PRODUCTS AND SERVICES` → table `pan20`, code column `CODE7`

**Excerpt of the filled table (RICA → FADN):**

```{r variable-pattern-preview, echo=FALSE, eval=TRUE}

# Load pre-filled dictionary
dict_conversion <- h_read_dictionary("dictionary_conversion_filled.xlsx")

dict_conversion$variable_pattern |> gt()

```

### Category pattern table

Each row maps one **national product code** to one **FADN product code**.
When several national codes map to the same FADN code, they will be **summed**
during aggregation (many-to-one). When one national code maps to several FADN
codes, an **allocation key** is needed (one-to-many).

| Column | Description | Example |
|---|---|---|
| `FADN_code_letter` | EU FADN product code | `CWHTC` |
| `national_code` | National product code (semicolon-separated if multiple) | `111` |
| `Group` | `CROPS`, `LIVESTOCK`, or `ANIMAL PRODUCTS AND SERVICES` | `CROPS` |

**Excerpt (from the provided data):**

```{r category-pattern-preview, echo=FALSE, eval=TRUE}

dict_conversion$category_pattern |> gt()

```

> **Note:** When multiple `national_code` rows share the same `FADN_code_letter`,
> the corresponding pivoted columns are summed. For example, RICA codes `115`
> and `116` both map to `CBRL` (barley), so `115_SUPER3 + 116_SUPER3` →
> `CBRL_TA`.

---

## Step-by-step workflow

### Step 0 – Load raw national data

If you have several

```{r load-raw}

data("mock_data")

RICA_20_raw <- list(
  rica = mock_data$rica_fict$rica,
  veg  = mock_data$rica_fict$veg,
  ani  = mock_data$rica_fict$ani,
  pan  = mock_data$rica_fict$pan
  )

RICA_19_raw <- list(
  rica = mock_data$rica_fict$rica,
  veg  = mock_data$rica_fict$veg,
  ani  = mock_data$rica_fict$ani,
  pan  = mock_data$rica_fict$pan
  )

RICA_18_raw <- list(
  rica = mock_data$rica_fict$rica,
  veg  = mock_data$rica_fict$veg,
  ani  = mock_data$rica_fict$ani,
  pan  = mock_data$rica_fict$pan
  )

```


## Multi-year workflow

National FADN data typically arrive as separate files per year. The package
provides `stack_yearly_tables()` to bind them into a single set of tables
before pivoting.

#### Step 0.1 — Load each year's raw data

```{r load-multi-year}
# Suppose you have raw RICA data loaded per year:
# RICA_18_raw, RICA_19_raw, RICA_20_raw — each a named list of data frames
# with elements like "rica", "veg", "ani", "pan", "tvn", "res", ...

yearly_data <- list(
  "2018" = RICA_18_raw,
  "2019" = RICA_19_raw,
  "2020" = RICA_20_raw
)
```

#### Step 0.2 — Stack into combined tables

```{r stack}
raw_tables <- h_stack_yearly_tables(yearly_data, year_col = "YEAR")
# Stacked 6 tables across 3 years.
#   rica: 22500 rows (3 years)
#   veg:  185000 rows (3 years)
#   ani:  67000 rows (3 years)
#   ...
```

Each table in `raw_tables` now has a `YEAR` column. You simply include it in
`id_cols` so that each farm × year combination stays on its own row.


### Notes on multi-year stacking

- **Column consistency**: `stack_yearly_tables()` uses `dplyr::bind_rows()`,
  which handles columns that appear in some years but not others (filled with
  `NA`). If a variable was renamed between survey years, harmonise the column
  names *before* stacking.

- **Year-specific unit changes**: If units changed between years (e.g., areas
  switched from ares to hectares in 2019), add year-conditional logic to the
  `unit_conversion` sheet of your dictionary, or pre-process the affected
  tables before stacking.

- **Table naming**: When stacking, the table names in the combined list must
  match those referenced in `table_config`. If your yearly files use
  year-suffixed names (e.g., `veg18`, `veg19`), rename them to a common name
  (e.g., `veg`) in each year's list before passing to `stack_yearly_tables()`.

```{r rename-example}
# Example: standardise table names before stacking
RICA_18_raw_std <- stats::setNames(
  RICA_18_raw,
  gsub("18$", "", names(RICA_18_raw))
)
# "veg18" becomes "veg", "ani18" becomes "ani", etc.
```

### Step 1 – Generate the template

The template pre-fills the structure with EU FADN variable suffixes and
category codes, and lists the national variable names found in your data.

```{r create-template}

h_create_dictionary_template(
  output_path = "dictionary_conversion.xlsx",
  raw_tables  = raw_tables,
  id_cols = "IDENT",
  table_config = list(
    CROPS = list(table = "veg", code_col = "CODE3"),
    LIVESTOCK = list(table = "ani", code_col = "CODE6"),
    `ANIMAL PRODUCTS AND SERVICES` = list(table = "pan", code_col = "CODE7")
  ),
  data_extra        = data_extra,
  national_code_col = "RICA_code_number",
  overwrite         = T
)

```

This creates an Excel workbook with the following sheets:

| Sheet | Content | Action required |
|---|---|---|
| `instructions` | Guidance | Read |
| `variable_pattern` | Pre-filled FADN suffixes + empty national suffix columns | **Fill `pattern_national` column** |
| `category_pattern` | Pre-filled from `data_extra` | **Review & complete** |
| `manual_matches` | For farm-level variables not following the pattern | **Fill** |
| `fadn_reference` | All EU FADN variable codes | Read-only reference |
| `category_reference` | All EU FADN category codes | Read-only reference |

### Step 2 – Fill the `variable_pattern` sheet

Open the Excel file and fill the `pattern_national` and `description_national` columns.
For each row, identify which column in the national long table corresponds to
the EU FADN suffix.

**Rules:**

- If no national equivalent exists, leave `pattern_national` empty (the variable
  will be `NA` in the output).
- If a national variable has no FADN equivalent but you want to keep it, add a
  row with `pattern_FADN` empty.
- The `Group` column must match exactly: `CROPS`, `LIVESTOCK`, or
  `ANIMAL PRODUCTS AND SERVICES`.

The table you fill should look like the excerpt shown above.

### Step 3 – Review the `category_pattern` sheet

This sheet is pre-filled from `data_extra`. Verify that:

- All national codes present in your data are listed
- Many-to-one mappings are correct (multiple national codes → one FADN code)
- For one-to-many cases, add an `allocation_key` column with proportions that
  sum to 1 across the FADN codes sharing the same national code

### Step 4 – Fill the `manual_matches` sheet (farm-level variables)

Variables from farm-level tables (`rica`, `res`, `tvn`) that don't follow
the `<CATEGORY>_<SUFFIX>` pattern must be matched here:

| fadn_variable | national_variable | fadn_table | national_table | conversion_factor | notes |
|---|---|---|---|---|---|
| `SE025` | `SUBUT`  | `STANDARD RESULTS` | `rica` | `0.01`  | UAA (ares → ha) |
| `SE131` | `PBRTTO` | `STANDARD RESULTS` | `rica` | `1`     | Total output |
| `SE290` | `CONSON` | `STANDARD RESULTS` | `tvn`  | `0.001` | N fertiliser (kg → t) |

Here is a pre-filled sheet for French FADN:

```{r manual_matches}

dict_conversion$manual_matches |> gt()

```


### Step 5 – Validate

```{r validate}

issues <- h_validate_dictionary(
  dictionary_path = "dictionary_conversion_filled.xlsx",
  raw_tables      = raw_tables,
  id_cols = "IDENT",
  table_config    = list(
    CROPS = list(table = "veg", code_col = "CODE3"),
    LIVESTOCK = list(table = "ani", code_col = "CODE6"),
    `ANIMAL PRODUCTS AND SERVICES` = list(table = "pan", code_col = "CODE7")
  ),
  verbose = TRUE
)

```

### Step 6 – Load the dictionary and convert

```{r convert}
# Load
dict_conversion <- h_read_dictionary("dictionary_conversion_filled.xlsx")

# Convert (pivot + rename + aggregate + unit conversion)
RICA_2020 <- h_pivot_national_fadn(
  raw_tables      = raw_tables,
  dict_conversion = dict_conversion,
  id_cols = c("IDENT","YEAR"),
  table_config    = list(
    CROPS = list(table = "veg", code_col = "CODE3"),
    LIVESTOCK = list(table = "ani", code_col = "CODE6"),
    `ANIMAL PRODUCTS AND SERVICES` = list(table = "pan", code_col = "CODE7")
  )
)

```

---

#### How the pivot works internally

The function `pivot_national_fadn()` performs the following for each product
group (using CROPS / `veg` as example):

```{r pivot-internals}
# --- Internal logic (pseudo-code, using CROPS as example) ---

# 1. Read which national columns to pivot from variable_pattern
#    For CROPS: SUPER3, SUPIR3, QPROD3, QVENT3, VAUTO3, VINT, VPBRT3,
#               VSTOD3, VSTOF3, VVENT3
crop_suffixes <- dict_conversion$variable_pattern %>%
  filter(Group == "CROPS", !is.na(pattern_national)) %>%
  select(pattern_national, pattern_FADN, unit_conversion)

# 2. Read category mapping (many-to-one allowed)
#    111 -> CWHTC, 115 -> CBRL, 116 -> CBRL, ...
crop_categories <- dict_conversion$category_pattern %>%
  filter(Group == "CROPS") %>%
  select(national_code, FADN_code_letter, allocation_key)

# 3. Identify numeric value columns, then summarise by farm × year × crop
#    This collapses production modes or duplicate rows within a farm-crop
group_cols  <- c("IDENT", "YEAR", "CODE3")
value_cols  <- setdiff(
  names(raw_tables$veg %>% select(where(is.numeric))),
  group_cols
)

veg_summed <- raw_tables$veg %>%
  summarise(
    across(all_of(value_cols), ~ sum(.x, na.rm = TRUE)),
    .by = all_of(group_cols)
  )

# 4. Pivot LONGER: one row per farm × year × crop × variable
#    This is the key difference from a direct pivot_wider approach —
#    going long first allows dictionary-driven joins
veg_long <- veg_summed %>%
  pivot_longer(
    cols      = all_of(value_cols),
    names_to  = "pattern_national",
    values_to = "value"
  ) %>%
  mutate(CODE3 = as.character(CODE3))
# Columns: IDENT, YEAR, CODE3, pattern_national, value
# e.g.:    FARM_0001, 2020, 111, SUPER3, 12.5

# 5. Join the FADN prefix (category mapping)
#    CODE3 "111" → FADN_code_letter "CWHTC", allocation_key 1
#    CODE3 "115" → FADN_code_letter "CBRL",  allocation_key 1
#    CODE3 "116" → FADN_code_letter "CBRL",  allocation_key 1
veg_long <- veg_long %>%
  left_join(crop_categories, by = c("CODE3" = "national_code"))

# 6. Join the FADN suffix (variable mapping) + unit conversion
#    pattern_national "SUPER3" → pattern_FADN "TA", unit_conversion 0.01
#    pattern_national "QPROD3" → pattern_FADN "PRQ", unit_conversion 0.01
veg_long <- veg_long %>%
  left_join(crop_suffixes, by = "pattern_national")

# 7. Aggregate: sum value × allocation_key × unit_conversion
#    This is where many-to-one categories are collapsed:
#    CBRL_TA = sum of (115_SUPER3 × 1 × 0.01) + (116_SUPER3 × 1 × 0.01)
#    Allocation keys allow splitting shared categories (e.g., eggs)
veg_summed_fadn <- veg_long %>%
  filter(!is.na(pattern_FADN), !is.na(FADN_code_letter)) %>%
  summarise(
    value = sum(value * allocation_key * unit_conversion, na.rm = TRUE),
    .by = c("IDENT", "YEAR", "FADN_code_letter", "pattern_FADN")
  )
# e.g.: FARM_0001, 2020, CWHTC, TA, 0.125
#        FARM_0001, 2020, CBRL,  TA, 0.340  (sum of 115 + 116)

# 8. Pivot WIDER: one row per farm-year, FADN-named columns
veg_final <- veg_summed_fadn %>%
  pivot_wider(
    id_cols     = c("IDENT", "YEAR"),
    names_from  = c("FADN_code_letter", "pattern_FADN"),
    values_from = value,
    values_fill = NA_real_
  )
# Columns: IDENT, YEAR, CWHTC_TA, CWHTC_PRQ, CBRL_TA, CBRL_PRQ, ...

```

#### Manual matches: farm-level variables

After all product groups are pivoted and joined, farm-level variables from
the main national table are added via the manual_matches dictionary sheet.
The function detects the column type automatically:

```{r}

# --- Manual match logic (pseudo-code) ---

# Character/factor columns (e.g., REGIO → NUTS2, OTEXE → TF14)
#   → take first non-NA value per farm-year (no numeric conversion)
raw_tables$rica %>%
  summarise(NUTS2 = first(REGIO[!is.na(REGIO)]), .by = c(IDENT, YEAR))

# Numeric columns (e.g., SAUTI → SE025, UTANSP → SE010)
#   → sum with unit conversion factor
raw_tables$rica %>%
  summarise(SE025 = sum(SAUTI * 1, na.rm = TRUE), .by = c(IDENT, YEAR))

# Redundant id mappings (e.g., IDENT → ID when IDENT is in id_cols)
#   → automatically skipped to avoid nonsensical aggregation

```

#### Summary of the pivot strategy

| Step | Operation | Purpose |
|------|-----------|---------|
| 1–2 | Read dictionary sheets | Load suffix map, category map, unit conversions |
| 3 | `summarise(.by = farm × crop)` | Collapse duplicate rows (e.g., production modes) |
| 4 | `pivot_longer()` | One row per farm × crop × variable (join-ready) |
| 5 | `left_join(categories)` | Attach FADN prefix + allocation key |
| 6 | `left_join(suffixes)` | Attach FADN suffix + unit conversion |
| 7 | `summarise(value × key × conv)` | Aggregate many-to-one categories |
| 8 | `pivot_wider()` | Final wide format: `CWHTC_TA`, `CBRL_PRQ`, … |
| 9 | Manual matches | Add farm-level character & numeric variables |

---

## Handling special cases

### Many national codes → one FADN code (aggregation)

Multiple RICA codes mapping to the same FADN code are **summed** automatically.

```
  RICA          FADN
  115 (winter barley)  ─┐
  116 (spring barley)  ─┼──► CBRL  (barley)
  117 (escourgeon)     ─┘
```

The pivoted columns `115_SUPER3`, `116_SUPER3`, `117_SUPER3` are summed into
`CBRL_TA`.

### One national code → many FADN codes (allocation)

When one RICA code must be split across multiple FADN codes, add an
`allocation_key` in the `category_pattern` sheet:

| FADN_code_letter | national_code | Group | allocation_key |
|---|---|---|---|
| `PEGGCONS` | `76` | ANIMAL PRODUCTS AND SERVICES | `0.95` |
| `PEGGHATCH` | `76` | ANIMAL PRODUCTS AND SERVICES | `0.05` |

The pivoted value for code `76` will be multiplied by the allocation key.

### RICA-specific unit quirks (livestock)

Some RICA variables encode units differently depending on the product code.
These are handled via **code-specific conversion factors** declared in the
`category_pattern` sheet or via a dedicated `unit_conversion` sheet:

| national_code | pattern_national | conversion_factor | notes |
|---|---|---|---|
| `911:965` | `EFFEC6` | `0.1` | tenths of head → heads |
| `974;975;983;984` | `EFFEC6` | `100000` | ten-thousandths → heads |
| `993` | `EFFEC6` | `10` | tens of hives → hives |
| `974;975;983;984` | `QPBRT6` | `1000` | thousands → units |

These are applied **before** pivoting in `pivot_national_fadn()`.

### FADN suffixes with no RICA equivalent

Rows in `variable_pattern` where `pattern_national` is empty (e.g., `A`, `ENA`,
`GM`, `ENTA`, `IRA`, `ALU`, `PI`, `SSN`, `SSV`, `SRN`, `SRV`, `NO`, `FCN`,
`FCV`, `FUN`, `FUV` for livestock) will produce `NA` columns in the output.
These can be computed downstream if data is available from other sources.

---

## Tips

- **Start with `variable_pattern`**: even a dozen suffix rows generate hundreds
  of variable mappings when crossed with the category codes.
- **The `Group` column is the join key** between the two tables. It must match
  exactly: `CROPS`, `LIVESTOCK`, or `ANIMAL PRODUCTS AND SERVICES`.
- **Version-control your dictionary**: the Excel file is the single source of
  truth. Track it alongside your code.
- **Re-use across years**: once built for one year, only minor updates are
  usually needed for subsequent years.
- **Use `validate_dictionary()` after every edit**: it catches missing mappings,
  invalid conversion factors, and orphaned category codes before they become
  silent `NA` in your results.


