---
title: "Inferring Agricultural Practices from FADN"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inferring Agricultural Practices from FADN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
link-citations: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
library(FADN2Footprint)
library(dplyr)
library(tidyr)
library(readxl)
library(gt)
```

# Why inference is needed (FADN doesn't record practices directly)

Our objective is to estimate the impact of farming practices in terms of
greenhouse gas (GHG) emissions and biodiversity impact. This requires
first to characterize farming practices implemented at the farm level
using data on the area of crop production, the quantity and type of
inputs used, the number of livestock, the amount and type of feed, and
the quantity and value of production ([Method
article](articles/practice-inference.html#sec-method-practice-intro) for
FADN primary variables used to characterize practices).

The **FADN2Footprint package** infers up to nine agricultural practices
from both FADN and LPIS variables: four within-field practices (tillage,
quantity and type of nitrogen fertilization, pesticide use) and five
landscape variables (livestock density, hedge density, mean field size,
ground cover, and crop diversity). Note that hedge density, mean field
size and ground cover cannot be inferred based solely on FADN.

Besides, FADN2Footprint also infer herding practices: feed (e.g., dry
matter intake and nitrogen excretion), rearing parameters (e.g., first
calving age, culling age).

| Data source / Practices | FADN<br>(+ Farm practices survey & feed tables for allocation keys)           | \+ LPIS & SENTINEL-2<br>+ parameters from national GHGIs |
|-------------------------|-------------------------------------------------------------------------------|----------------------------------------------------------|
| Crop practices          | •Fertilizer quantity and quality •Pesticides •Tillage •Crop diversity •Yields | •Hedge density •Mean field size •Ground cover            |
| Grassland practices     | •Fertilizer quantity and quality •Livestock density                           | •Hedge density •Yields                                   |
| Herding practices       | •Feed intake •Rearing parameters •Production / Yields                         | •Manure management systems •Animal weight                |

To infer all these practices, install the package, build a S4
FADN2Footprint object and run the `infer_practices` function:

```{r}

# Install the package
#install.packages('FADN2Footprint')

# Load your data (here we use the mock data and dictionary included in the package)
data("mock_data", package = "FADN2Footprint")
data("dict_FADN", package = "FADN2Footprint")

fict_data <- mock_data$fadn_fict

# Build the S4 object
my_object = FADN2Footprint::data_4FADN2Footprint(
  df = fict_data,
  var_dict = dict_FADN,
  id_cols = c("ID","YEAR","COUNTRY")
)

# Infer the practices
my_object_w_practices <- FADN2Footprint::infer_practices(my_object)

```

```{r tmp}

load("C:/Users/srhuet/OneDrive/Research/GitHub/Organic_prod_in_Europe/data_raw/FADN_16_18_object_w_practices.RData")

my_object_w_practices = FADN_16_18_GHGE

```

# Crop-side practices

## Tillage

Tillage intensity (L diesel/ha) is estimated from off-road diesel
consumption, allocated to crops by area ([Method
article](articles/practice-inference.html#sec-method-practices-tillage)).XX
le lien ne fonctionne pas

## Nitrogen fertilization

Nitrogen input is estimated separately for mineral and organic
fertilizers ([Method
article](articles/practice-inference.html#sec-method-practices-Nferti)).

For mineral fertilizers, the total farm-level amount (reported in FADN)
is allocated to the different crops using the French average mineral
nitrogen input per crop as a distribution key (Ministère De
L’Agriculture - SSP, 2019).

For organic fertilizers, since FADN data on organic nitrogen is
incomplete, estimates depend on livestock presence. For farms without
livestock, a standard national average of organic nitrogen input per
crop is assigned (Ministère De L’Agriculture - SSP, 2019). For farms
with livestock, we assume all excreted nitrogen is spread on-farm, which
is estimated in accordance with IPCC recommendations (Eggleston et al.,
2006; Calvo Buendia et al., 2019; [Method
article](articles/practice-inference.html#sec-method-practices-Nferti)).
Nitrogen is then allocated across crops using the French average organic
nitrogen input per crop as a distribution key (Ministère De
L’Agriculture - SSP, 2019).

Let's have a look at the inferred fertilisation practice. What does it
say for common wheat:

```{r practice_Nferti_wheat, message=FALSE}

my_Nferti_wheat <- my_object_w_practices@practices$crops$N_ferti |>
  # filter for wheat
  dplyr::filter(FADN_code_letter == "CWHTC")

library(ggplot2)
library(dplyr)
library(tidyr)

# remove outliers
outlier_thrld = quantile(my_Nferti_wheat$N_ferti,0.95,na.rm = T)

# Reshape data to long format for easier plotting
my_Nferti_wheat_long <- my_Nferti_wheat |>
  dplyr::filter(N_ferti <= outlier_thrld) |>
  select(ID, YEAR, COUNTRY, ORGANIC, N_ferti, N_ferti_min, N_ferti_org) %>%
  pivot_longer(
    cols = c(N_ferti, N_ferti_min, N_ferti_org),
    names_to = "variable",
    values_to = "value_kgNha"
  ) %>%
  mutate(
    variable = recode(variable,
                      "N_ferti"     = "Total N",
                      "N_ferti_min" = "Mineral N",
                      "N_ferti_org" = "Organic N"
    ),
    variable = factor(variable,
                      levels = c("Total N", "Mineral N", "Organic N")),
    ORGANIC = as.character(ORGANIC),
    ORGANIC = recode(ORGANIC,
                     'TRUE' =  "Organic farming",
                     'FALSE' = "Non-Organic farming"
    ),
    ORGANIC = factor(ORGANIC, 
                     levels = c("Organic farming", "Non-Organic farming"))
    
  )

# Plot
ggplot(my_Nferti_wheat_long) +
  aes(x = COUNTRY, y = value_kgNha, fill = ORGANIC) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8, color = "grey30") +
  coord_flip() +
  facet_wrap(vars(variable), scales = "free") +
  scale_fill_manual(values = c(
    "Organic farming" = "#4CAF50",
    "Non-Organic farming" = "#2196F3"
  )) +
  labs(
    title    = "Nitrogen Fertilization of Wheat",
    subtitle = "Distribution by country and N source",
    x        = NULL,
    y        = "N application (kg N/ha)",
    fill     = "Organic farming",
    caption  = paste0("Outliers above ",names(outlier_thrld)," excluded for readability")
  ) +
  theme_minimal()+
  theme(
    legend.position  = "bottom",
    strip.background = element_rect(fill = "grey90"),
    strip.text       = element_text(face = "bold"),
    axis.text.x      = element_text(angle = 20, hjust = 1),
    plot.title       = element_text(face = "bold")
  )


```

## Pesticide use {#sec-pesticides}

Pesticide intensity is estimated using FADN pesticide expenditure (€),
distributed across crops based on the national Treatment Frequency Index
(TFI; Ministère De L’Agriculture - SSP, 2019). The indicator is divided
by 3.8 in organic farms to account for the lower unit toxicity of
mineral pesticides ([Method
article](articles/practice-inference.html#sec-method-practices-pesticides)).

## Yield (observed vs. imputed)

Yield in FADN2Footprint are obtained directly from FADN when both crop
area and production are registered. When FADN production is missing, we
impute default average yields. For forage/grass, where FADN often
records area but not production, we impute yields from external
reference data (French forage yields aggregated by crop type and NUTS3;
Agreste SAA 2020; `data_extra$yield_SAA_Agreste_2020`), joined at a
NUTS3 level (when available, at country level when not) and used to
infer the missing production. For missing production of others crops, we
impute the user data set average of the same crop at the NUTS2 level if
available, or at the whole data set level otherwise. For some crops
where FADN record no production, we impute average of similar crops
based on the following matching table:

| **FADN code** | **Description**                           | **Crop average used**                                                                |
|---------------|-------------------------------------------|--------------------------------------------------------------------------------------|
| CARAOTH       | Other arable land crops                   | Average yield of other arable crops (code 10110 to 10690; see `data_extra$crops`)    |
| CCRPPERMOTH   | Other permanent crops                     | Average yield of other permanent crops (code 40111 to 40500; see `data_extra$crops`) |
| CPERMUG       | Permanent crops under glass               | Average yield of other permanent crops (code 40111 to 40500; see `data_extra$crops`) |
| CFIBOTH       | Other fiber plants                        | Average yield of other fiber crops (code 10609 and 10610; see `data_extra$crops`)    |
| CKGAR         | Kitchen gardens                           | Average of vegetables (code 10711 to 10790; see `data_extra$crops`)                  |
| CFLW          | Flowering and ornamental plants           | Set to zero                                                                          |
| CFLWB         | Flower bulbs, corms and tubers            | Set to zero                                                                          |
| CFLWCUT       | Cut flowers and flower buds               | Set to zero                                                                          |
| CFLWOUT       | Flowers and ornamental plants Outdoor     | Set to zero                                                                          |
| CFLWUG        | Flowers and ornamental plants Under glass | Set to zero                                                                          |
| CNURS         | Nurseries                                 | Set to zero                                                                          |
| CLNDOTH       | Other land                                | Set to zero as it is not arable nor grass lands                                      |

# Landscape variables

We estimate four landscape variables ([Method
article](articles/practice-inference.html#sec-method-practices-landscape-var)).
Hedge density is the total hedge length divided the utilized
agricultural area (UAA) of the farm ([Method
article](articles/practice-inference.html#sec-method-practices-hedge)).
Mean field size is the UAA divided by the number of plots, based on LPIS
data ([Method
article](articles/practice-inference.html#sec-method-practices-mean-field-size)).
Crop diversity is the Shannon diversity index derived from the FADN data
([Method
article](articles/practice-inference.html#sec-method-practices-crop-div)).
Ground cover is the average number of days per year without vegetation,
estimated from Sentinel-2 raster data ([Method
article](articles/practice-inference.html#sec-method-practices-ground-cover)).

Within a farm, all crops are associated to the same landscape variable
(e.g. hedge density in m/ha). This contrasts with within field practices
(fertilization, pesticides use, tillage, see allocation keys above). A
different hedge density is associated with permanent grassland (see
below).

# Practices in grasslands

For grasslands, we consider four agricultural practices: hedge density,
total nitrogen fertilization as well as the share of mineral fertilizer
and livestock density, estimated as the number of grazing animals (e.g.,
cattle, sheep, horses) per hectare of main forage area (e.g., grass,
green maize). Indeed, permanent grasslands are not ploughed, do not
receive pesticides, and are not associated with any crop diversity. We
also considered negligible the effect of mean field size on biodiversity
in grasslands as few studies are available and report contradictory
effects (Dumont et al., 2016; Lescourret et al., 2025)). Permanent
grasslands also have a year-round ground cover. Note that temporary
grasslands are considered as crops as they are part of a crop rotation.

```{r}

library(ggplot2)
library(dplyr)
library(patchwork) # For combining the plots
library(scales)

# --- 1. Grassland Area per Farm by Country ---
# We use a log10 scale because FADN farm areas are highly skewed
plot_area <- my_data |>
  filter(!is.na(area_ha), area_ha > 0) |>
  # Order countries by median area for a cleaner look
  mutate(COUNTRY = reorder(COUNTRY, area_ha, FUN = median, na.rm = TRUE)) |>
  ggplot(aes(x = COUNTRY, y = area_ha)) +
  geom_boxplot(fill = "#81C784", colour = "#2E7D32", outlier.alpha = 0.2, outlier.size = 1) +
  scale_y_log10(
    labels = scales::comma_format(),
    breaks = c(1, 10, 100, 1000)
  ) +
  coord_flip() +
  labs(
    title = "Grassland Area per Farm",
    subtitle = "Logarithmic scale applied",
    x = NULL,
    y = "Area (ha) [log10]"
  ) +
  theme_light(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey40"),
    panel.grid.minor = element_blank()
  )

# --- 2. Grassland Productivity (Yield) by Country ---
# We cap the maximum y-value at the 99th percentile to hide extreme outliers
yield_cap <- quantile(my_data$yield, 0.99, na.rm = TRUE)

plot_yield <- my_data |>
  filter(!is.na(yield), yield > 0) |>
  # Order countries by median yield
  mutate(COUNTRY = reorder(COUNTRY, yield, FUN = median, na.rm = TRUE)) |>
  ggplot(aes(x = COUNTRY, y = yield)) +
  geom_boxplot(fill = "#64B5F6", colour = "#1565C0", outlier.alpha = 0.2, outlier.size = 1) +
  scale_y_continuous(
    limits = c(0, yield_cap), 
    labels = scales::comma_format()
  ) +
  coord_flip() +
  labs(
    title = "Grassland Productivity",
    subtitle = "Yield distribution by country",
    x = NULL,
    y = "Yield (t/ha)",
    caption = "Yield outliers above the 99th percentile are excluded"
  ) +
  theme_light(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(colour = "grey40"),
    plot.caption  = element_text(colour = "grey50", size = 8, hjust = 1),
    panel.grid.minor = element_blank()
  )

# --- Combine plots side-by-side using patchwork ---
combined_plot <- plot_area + plot_yield + 
  plot_annotation(
    title = "FADN Grassland Overview",
    theme = theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
  )

combined_plot


```

# Herding practices

## Herd structure

Following the JRC ILCD handbook (JRC, 2010), we avoid allocation
wherever possible by virtual subdivision of processes. In particular, we
divide farms into a dairy activity (with all animals required to sustain
observed milk production, including calves and heifers raised to renew
dairy cows) and a meat activity (with all animals dedicated to meat,
including calves and suckling cows necessary to renew fattening beefs).
Based on FADN data, we reconstruct the cattle herd structure and model
the rearing process by estimating animal flows between categories
(inflows and outflows). From these flows, we calculate key rearing
parameters, specifically residence time for each livestock category,
offspring rates per breeder, and the age at first calving. Then, we
calculate the number of animals involved in the dairy activity by
partitioning the number of animals in categories upstream of dairy cows
based on the estimated rearing parameters ([Method
article](articles/practice-inference.html#sec-method-practices-rearing-param)).

## Feed

We estimate the quantity and quality of the livestock feed, considering
both on-farm (i.e., farm-grown) and off-farm (i.e., purchased) feed to
form a "pseudo-farm" ([Method
article](articles/practice-inference.html#sec-method-practices-feed)).

On-farm produced feed quantity is calculated as the recorded crop
production minus sales, with grass yield - not provided in FADN -
assumed equal to the NUTS2 average (Agreste, 2020). This net available
on-farm biomass is then allocated to the farm's livestock categories
proportional to their theoretical dry matter requirements for each crop.

Purchased feed quantity is estimated based on FADN expenditures
variables for concentrated feedstuffs and coarse fodder (Section
A.2.5.2). These expenditures are allocated to livestock species in
proportion to their theoretical ration cost for each feedstuff
(concentrates and roughage). The observed FADN expenditure is then
allocated to specific feed types (e.g., wheat, barley, maize, soybean
meal) according to these theoretical value shares. Finally, these
allocated values per feed type and species are converted into physical
quantities (tonnes) by dividing by the corresponding Eurostat unit
prices.

This estimation relies on a theoretical feed ration constructed for each
livestock species. Theoretical rations are determined by combining the
total dry matter intake requirements per livestock unit (derived from
the AROPAJ model; Jayet et al., 2023) with the species-specific volumes
of each feed consumed nationally (Sailley et al., 2021; see
`data_extra$Sailley_2021_feed_flows`).

Feed quality parameters - gross energy (GE), crude protein content (CP),
and ash – are calculated by applying nutritional coefficients from feed
tables (Tran, 2002; [Method
article](articles/practice-inference.html#sec-method-practices-feed-qty-qlty))
to the estimated dry matter quantities. Feed digestibility is retrieved
from the methodoly of the New Zealand Ministry for Primary Industries
(*New Zealand Farm Emissions Method: A farm-level approach for
estimating biogenic emissions*, 2024). To exclude aberrant estimations,
we remove farms with extreme GE supplied to their main livestock
category. Specifically, we exclude farms with, for the livestock
category they have the highest number of livestock units, a GE exceeding
the mean + 3 × standard deviation (a common threshold in the
literature), below 0.25 × the mean GE for dairy cows (since even
low-producing dairy cows are unlikely to be sustained on less than a
quarter of the average energy intake), or below 0.10 x the mean GE for
non-dairy cow livestock categories (Table A4).

Then, we estimate the crop practices implemented to produce the feed,
and consider practices associated with both on-farm and purchased feed
([Method
article](articles/practice-inference.html#sec-method-practices-feed-qty-qlty)).
Farming practices associated with on-farm feed are obtained from
farm-specific data (Sectionto XX?). Practices associated with purchased
feed reflect in-study national crop averages, distinguishing organic vs
conventional, except for soybean meal. The soybean meals are assumed to
be imported from Brazil (Overmars et al., 20xx) and asociated practices
are retrieved from Lindner et al. (2022).

## Manure management system

The fraction of manure handled by each manure management system (AWMS)
on a farm is not measured directly in FADN for most countries. Instead
AWMS fractions are inferred by combining the observed herd and farm
characteristics from FADN with country (and where available
climate‑region) AWMS allocations reported in UNFCCC CRF tables, and then
refined where farm‑level information exists (farm practices survey,
LPIS, housing/grazing proxies). Those inferred AWMS fractions are then
used together with IPCC/UNFCCC MCF and Bo parameters to compute CH4 and
(together with N excretion and EF) N2O from manure management.

```{r}
library(ggplot2)
library(dplyr)
library(scales)

# Extract and clean the data
awms_plot <- FADN2Footprint::UNFCCC_data$table3Bas2 |>
  # filter for dairy cows
  dplyr::filter(Animal_cat == "dairy_cattle") |>
  # (Optional) filter for a single year if the dataset contains multiple
  # filter(YEAR == "2018") |> 
  ggplot(aes(x = Country_ISO_3166_1_A3, y = AWMS, fill = MMS)) +
  # Use position = "fill" to normalize stacks to 100% (handles minor rounding variations)
  geom_col(position = "fill", width = 0.8) + 
  scale_y_continuous(labels = scales::percent_format()) +
  # Using a colorbrewer palette for clear categorical distinctions
  scale_fill_brewer(palette = "Set3") +
  coord_flip() +
  labs(
    title = "Share of Manure Management Systems by Country",
    subtitle = "Relative allocation for dairy cattle from UNFCCC",
    x = "Country",
    y = "Share of AWMS",
    fill = "Management System"
  ) +
  theme_light(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    legend.key.size = unit(0.5, "cm"),
    strip.background = element_rect(fill = "#e0e0e0"),
    strip.text = element_text(color = "black", face = "bold"),
    panel.grid.major.x = element_blank()
  )

awms_plot

```

## Animal weights

Animal weights are not reported in FADN. We use the national animal
weights reported in UNFCCC CRF tables.
