---
title: "Pairing FADN with ADEME"
author: "Sarah HUET"
date: "2026-07-06"
output: html_document
---

This script aim at pairing FADN farms to those in the ADEME data base, withou using the SIRET as we don't have it locally.


# Setup

```{r setup}

library(dplyr)
library(tidyr)
library(stringr)
library(gt)
library(ggplot2)

library(purrr)

# For distance-based matching
library(cluster)      # daisy() for Gower distance (mixed scales)
# library(clue)       # solve_LSAP() for optimal 1:1 assignment (optional)


devtools::load_all()
#library(FADN2Footprint)

```

#TODO: check cattle_meat VS meat

# Load data

```{r load_data, echo=FALSE, message=TRUE, warning=FALSE}

## Load raw data and run model
#source("validation_FADN_data.R")
#source("validation_FADN_data_offline.R")
## 
#load("../../data_raw/FADN_FRA_2018_2021_obj_practices.RData")
load("../../data_raw/FADN_FRA_2018_2021_obj_GHGE.RData")

cat("FADN dataset loaded:", nrow(my_object_GHGE@farm), "farm-year observations\n")
cat("Countries covered:", paste(unique(my_object_GHGE@farm$COUNTRY), collapse = ", "), "\n")
cat("Years covered:", paste(sort(unique(my_object_GHGE@farm$YEAR)), collapse = ", "), "\n")

```

```{r ADEME_data}

library(readr)
Df_Specific_ademe <- read_csv("C:/Users/srhuet/OneDrive/Research/GitHub/FADN2Footprint_validation/data/raw/Df_Specific_ademe.csv")



```

# Harmonize the ADEME database
The ADEME crops are very granular; we aggregate them into the broad land-use categories that FADN can reproduce, and we build a NUTS2 code from the French département.


```{r harmonize_data}


# --- 2.1 Map French département -> NUTS2 -------------------------------------

# NUTS2 and NUTS3 codes
library(readxl)
fr_nuts3 <- read_excel("C:/Users/srhuet/OneDrive/Research/Data/EUROSTAT/NUTS2024-NUTS2027.xlsx", 
                       sheet = "France")

ademe <- Df_Specific_ademe |>
     dplyr::mutate(dep_code = substr(exploitation_code_postal, start = 1, stop = 2),
          YEAR = as.character(year)
     ) |>
     dplyr::left_join(fr_nuts3, by = "dep_code")

# --- 2.2 Map type of farming -------------------------------------------------



# --- 2.2 Aggregate crop areas into FADN-comparable land-use groups -----------
ademe <- ademe %>%
     mutate(
          area_cereals = rowSums(across(c(assolement_ble_tendre, assolement_ble_dur,
                                          assolement_mais_grain, assolement_orge,
                                          assolement_autres_cereales)), na.rm = TRUE),
          area_oilseed = rowSums(across(c(assolement_colza_et_navette, assolement_tournesol,
                                          assolement_soja, assolement_autres_oleagineux)), na.rm = TRUE),
          area_protein = rowSums(across(c(assolement_pois_proteagineux,
                                          assolement_autres_proteagineux,
                                          assolement_legumes_secs_ou_a_grains)), na.rm = TRUE),
          area_forage  = rowSums(across(c(assolement_luzerne, assolement_mais_fourrage_et_ensilage,
                                          assolement_autres_fourrage_annuels,
                                          assolement_prairies_temporaires_legumineuses,
                                          assolement_prairies_temporaires_graminees,
                                          assolement_prairies_temporaires_melange_graminees_et_legumineuse)),
                                 na.rm = TRUE),
          area_permgrass = assolement_prairies_permanentes,
          area_permcrop  = rowSums(across(c(assolement_vignes, assolement_vergers,
                                            assolement_autres_cultures_permanentes)), na.rm = TRUE),
          area_veg       = rowSums(across(c(assolement_legumes_fleurs_fruit_hors_arboriculture_plein_air,
                                            assolement_legumes_fleurs_fruit_hors_arboriculture_sous_serres)),
                                   na.rm = TRUE),
          SAU = assolement_sau_totale,
          
          # Animal numbers (heads)
          n_dairy_cattle = cheptel_bovins_lait,
          n_beef_cattle  = cheptel_bovins_viande,
          n_sheep        = rowSums(across(c(cheptel_ovins_lait, cheptel_ovins_viande)), na.rm = TRUE),
          n_goats        = cheptel_caprins,
          n_pigs         = rowSums(across(c(cheptel_porcs_engraissement, cheptel_truies,
                                            cheptel_porcelets)), na.rm = TRUE),
          n_poultry      = rowSums(across(c(cheptel_poules_pondeuses, cheptel_volailles_de_chair,
                                            cheptel_palmipedes_gras)), na.rm = TRUE),
          
          # Type of farming: OTEX. Reduce to a coarse grouping comparable to FADN TF8/TF14.
          TF = otex_to_tf(exploitation_otex)   # <- define your own recode function
     ) %>%
     mutate(ademe_row = row_number())        # keep a stable id

```


Note: otex_to_tf() is a helper you must define mapping the ADEME OTEX code to the same
classification you use on the FADN side (TF8 or TF14). Matching quality depends heavily on
this being consistent between the two sources.


3. Harmonize the FADN side (your package output)
FADN crops/herds are in long format (one row per crop/species). We pivot them to the same wide, aggregated schema.
# --- 3.1 Farm-level table ----------------------------------------------------
fadn_farm <- my_object_GHGE@farm %>%
transmute(
ID, year = as.integer(YEAR), NUTS2,
TF = as.character(TF8),          # use same classification as ademe TF
SAU = SE025                      # total UAA (ha) -- adjust to the right variable
)

# --- 3.2 Aggregate crop areas into the same land-use groups ------------------
# 'land_use_type' / 'species' from the crop table drives the grouping.
fadn_crop_wide <- my_object_GHGE@crop %>%
mutate(group = recode_land_use(land_use_type, species)) %>%   # your helper
group_by(ID, YEAR) %>%
summarise(
area_cereals   = sum(area_ha[group == "cereals"],   na.rm = TRUE),
area_oilseed   = sum(area_ha[group == "oilseed"],   na.rm = TRUE),
area_protein   = sum(area_ha[group == "protein"],   na.rm = TRUE),
area_forage    = sum(area_ha[group == "forage"],    na.rm = TRUE),
area_permgrass = sum(area_ha[group == "permgrass"], na.rm = TRUE),
area_permcrop  = sum(area_ha[group == "permcrop"],  na.rm = TRUE),
area_veg       = sum(area_ha[group == "veg"],       na.rm = TRUE),
.groups = "drop"
) %>%
rename(year = YEAR)

# --- 3.3 Aggregate herd numbers into the same species groups -----------------
fadn_herd_wide <- my_object_GHGE@herd %>%
mutate(group = recode_species(species)) %>%                   # your helper
group_by(ID, YEAR) %>%
summarise(
n_dairy_cattle = sum(AN[group == "dairy_cattle"], na.rm = TRUE),
n_beef_cattle  = sum(AN[group == "beef_cattle"],  na.rm = TRUE),
n_sheep        = sum(AN[group == "sheep"],        na.rm = TRUE),
n_goats        = sum(AN[group == "goats"],        na.rm = TRUE),
n_pigs         = sum(AN[group == "pigs"],         na.rm = TRUE),
n_poultry      = sum(AN[group == "poultry"],      na.rm = TRUE),
.groups = "drop"
) %>%
rename(year = YEAR)

# --- 3.4 Combine into one wide FADN profile ---------------------------------
fadn <- fadn_farm %>%
left_join(fadn_crop_wide, by = c("ID", "year")) %>%
left_join(fadn_herd_wide, by = c("ID", "year")) %>%
mutate(across(starts_with("area_"), ~replace_na(., 0)),
across(starts_with("n_"),    ~replace_na(., 0))) %>%
mutate(fadn_row = row_number())

4. Define the matching variables and a scaling function
match_vars <- c("area_cereals","area_oilseed","area_protein","area_forage",
"area_permgrass","area_permcrop","area_veg","SAU",
"n_dairy_cattle","n_beef_cattle","n_sheep","n_goats",
"n_pigs","n_poultry")

block_vars <- c("NUTS2", "TF", "year")   # exact-match blocking keys

5. The matching engine (blocking + Gower nearest neighbour)
#' Match one ADEME record to the nearest FADN record within its block.
#'
#' @param ademe_df harmonised ADEME data
#' @param fadn_df  harmonised FADN data
#' @param match_vars numeric variables used for the distance
#' @param block_vars exact-match keys
#' @param max_dist   maximum acceptable Gower distance (0-1); NA above threshold
match_ademe_fadn <- function(ademe_df, fadn_df,
match_vars, block_vars,
max_dist = 0.15) {

# build a block key
ademe_df$.block <- do.call(paste, c(ademe_df[block_vars], sep = "|"))
fadn_df$.block  <- do.call(paste, c(fadn_df[block_vars],  sep = "|"))

common_blocks <- intersect(unique(ademe_df$.block), unique(fadn_df$.block))

# scaling ranges computed globally so distances are comparable across blocks
ranges <- sapply(match_vars, function(v) {
diff(range(c(ademe_df[[v]], fadn_df[[v]]), na.rm = TRUE))
})
ranges[ranges == 0] <- 1

results <- map_dfr(common_blocks, function(bk) {
a <- ademe_df[ademe_df$.block == bk, ]
f <- fadn_df[fadn_df$.block == bk, ]

# matrix of scaled values
A <- scale(as.matrix(a[match_vars]), center = FALSE, scale = ranges)
F <- scale(as.matrix(f[match_vars]), center = FALSE, scale = ranges)

# pairwise Manhattan distance / n vars  -> Gower-like distance in [0,1]
# (fast, no need for daisy on the full set)
d <- outer(seq_len(nrow(A)), seq_len(nrow(F)), Vectorize(function(i, j) {
mean(abs(A[i, ] - F[j, ]))
}))

nn_idx  <- apply(d, 1, which.min)
nn_dist <- apply(d, 1, min)

tibble(
ademe_row = a$ademe_row,
fadn_row  = f$fadn_row[nn_idx],
fadn_ID   = f$ID[nn_idx],
block     = bk,
dist      = nn_dist
) %>%
mutate(fadn_row = ifelse(dist <= max_dist, fadn_row, NA),
fadn_ID  = ifelse(dist <= max_dist, fadn_ID,  NA))
})

results
}

matches <- match_ademe_fadn(ademe, fadn, match_vars, block_vars, max_dist = 0.15)

6. Assemble the paired dataset
paired <- matches %>%
filter(!is.na(fadn_row)) %>%
left_join(ademe, by = "ademe_row",
suffix = c("", "")) %>%
left_join(fadn %>% rename_with(~paste0("fadn_", .x), -fadn_row),
by = "fadn_row")

# Quick diagnostic: how many ADEME records were matched?
cat(sprintf("Matched %d of %d ADEME records (%.1f%%)\n",
sum(!is.na(matches$fadn_row)), nrow(ademe),
100 * mean(!is.na(matches$fadn_row))))

# Distribution of matching distances
summary(matches$dist)

7. Compare the GHG estimates
comparison <- paired %>%
transmute(
fadn_ID, NUTS2, TF, year, dist,
ademe_ghg = emissions_ges_totales_exploitation,
# your package GHG total: join it from the relevant slot of my_object_GHGE
)

# e.g. plot
library(ggplot2)
ggplot(comparison, aes(ademe_ghg, my_ghg)) +
geom_point(aes(colour = dist), alpha = .5) +
geom_abline(slope = 1, intercept = 0, linetype = 2) +
labs(x = "ADEME diagnosis (t CO2e)", y = "FADN2Footprint (t CO2e)")
