---
title: "FADN2Footprint — Validation Report"
subtitle: "Comparison of package estimates with literature values"
author: "Sarah HUET & Valentin BELLASSEN"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true        
  pdf:
    toc: true
execute:
  echo: true
  warning: false
editor: 
  markdown: 
    wrap: 150
bibliography: ../../vignettes/references.bib
link-citations: true
---

# Purpose

This report compares the outputs of `FADN2Footprint` (v`r packageVersion("FADN2Footprint")`)
against published literature ranges.
An estimate is flagged **in range** if it falls within the literature min–max,
**within tolerance** if the relative deviation from the literature midpoint is
below a per-indicator threshold, or **out of range** otherwise.

# Setup

```{r setup}

library(dplyr)
library(tidyr)
library(stringr)
library(gt)

library(ggplot2)

devtools::load_all()
#library(FADN2Footprint)

```

#TODO: check cattle_meat VS meat

# Produce package estimates

```{r load_data, echo=FALSE, message=TRUE, warning=FALSE}

## Load raw data and run model
#source("validation_FADN_data.R")
#source("validation_FADN_data_offline.R")
## 
#load("../../data_raw/FADN_FRA_2018_2021_obj_practices.RData")
load("../../data_raw/FADN_FRA_2018_2021_obj_GHGE.RData")
my_object_GHGE <- m_prune_farms(my_object_GHGE,
                                YEAR %in% c("2018", "2019", "2020"))

cat("FADN dataset loaded:", nrow(my_object_GHGE@farm), "farm-year observations\n")
cat("Countries covered:", paste(unique(my_object_GHGE@farm$COUNTRY), collapse = ", "), "\n")
cat("Years covered:", paste(sort(unique(my_object_GHGE@farm$YEAR)), collapse = ", "), "\n")

```

```{r summaries}


object = my_object_GHGE

# Crops
summary(object@crop)

(object@crop |> 
    dplyr::filter(is.na(prod_t)))

#sort(table(object@crop |> filter(is.na(prod_t)) |> pull(FADN_code_letter)), decreasing = TRUE)
#table(object@crop |> filter(is.na(prod_t)) |> pull(keep_sup5pc))
#
#
## GHGE crops
#
summary(object@footprints$GHGE$GHGE_crops)
#
object@footprints$GHGE$GHGE_crops |> 
  dplyr::filter(!is.finite(N2O_d_kgCO2e_per_t)) #|> View()
#sort(table(object@footprints$GHGE$GHGE_crops |> filter(is.na(prod_t)) |> pull(FADN_code_letter)), decreasing = TRUE)

```


# Literature references

```{r load_literature, message=FALSE}

# TOTAL GHGE ----

lit_GHGE_total <- readxl::read_excel(
  "../../inst/validation/validation_references.xlsx",
  sheet = "summary_total_GHGE",
  col_types = c("text", "numeric", "numeric", "text", "text")
) |>
  # filter for references with both per ha and per t
  dplyr::filter(GHGE_kgCO2e_kg >0 
                | GHGE_kgCO2e_ha >0) |>
  # filter for crops, meat, milk, and eggs 
  dplyr::filter(grepl("wheat|maize_grain|rapeseed|milk|meat|eggs", product)) |>
  # filter values for France if present or EU otherwise for each product
  dplyr::group_by(product) |>
  dplyr::filter(
    if (any(grepl("France", area))) grepl("France", area) else TRUE
  ) |>
  dplyr::ungroup() |>
  # add set
  dplyr::mutate(set = "GHG Emissions")

# COMPONENTS OF GHGE ----

lit_GHGE_components <- readxl::read_excel(
  "../../inst/validation/validation_references.xlsx",
  sheet = "summary_share_components"
) |>
  # filter for crops, meat and milk 
  dplyr::filter(grepl("wheat|maize_grain|rapeseed|milk|meat|eggs", product)) |>
  # filter values for France if present or EU otherwise for each product and component
  dplyr::group_by(product, Harmonized_Category) |>
  dplyr::filter(
    if (any(grepl("France", area))) grepl("France", area) else TRUE
  ) |>
  dplyr::ungroup() |>
  # remove unclassified components
  dplyr::filter(!is.na(Harmonized_Category)) |>
  # sum up by product,Harmonized_Category,sheet,area
  dplyr::summarise(
    dplyr::across(
      c(GHGE_kgCO2e_ha,GHGE_kgCO2e_kg),
      ~ sum(.x)
    ),
    .by = c('product','Harmonized_Category','sheet','area')
  ) |>
  # add set
  dplyr::mutate(set = "GHG Emissions")

cat("Literature (total GHGE) loaded:", nrow(lit_GHGE_total), "rows\n")
cat("Literature (components of GHGE) loaded:", nrow(lit_GHGE_components), "rows\n")

lit_agri_pract <- readxl::read_excel(
  "../../inst/validation/validation_references.xlsx",
  sheet = "summary_agri_practices"
)

# add EPK averages for wheat fertilization ----
tmp = data_extra$PKGC_N_ferti |>
  dplyr::filter(EPK_ferti_crop_name == "Ble_tendre" & ORGANIC == 0) |>
  dplyr::select(N_min,Nmin_sd_w,N_org,Norg_sd_w)

lit_agri_pract <- lit_agri_pract |>
  dplyr::bind_rows(
    tibble::tibble(
      indicator = c("mineral fertilization", "organic fertilizer"),
      unit = "kg N / ha",
      product = "wheat",
      area = "France",
      lit_midpoint = c(tmp$N_min, tmp$N_org),
      lit_min = 0,
      lit_max = c((tmp$N_min + tmp$Nmin_sd_w), (tmp$N_org + tmp$Norg_sd_w)),
      #tolerance_pct = "More than half\n&\nLess than twice",
      lit_source = "EPK 2017"
    )
  ) |>
  # add set
  dplyr::mutate(set = "Agricultural practices")


cat("Literature (agricultural practices) loaded:", nrow(lit_agri_pract), "rows\n")

lit_GHGE_total |> arrange(set,product,sheet) |> gt()
lit_GHGE_components |> arrange(set,product,Harmonized_Category,sheet) |> gt()
lit_agri_pract |> arrange(set,indicator) |> gt()

```


```{r lit_table, message=FALSE}

lit_table = dplyr::bind_rows(
  lit_agri_pract,
  lit_GHGE_total |>
    dplyr::mutate(GHGE_kgCO2e_t = GHGE_kgCO2e_kg*10^3) |>
    tidyr::pivot_longer(cols = c(GHGE_kgCO2e_ha,GHGE_kgCO2e_t),
                        names_to = "indicator") |>
    dplyr::filter(!is.na(value)) |>
    dplyr::summarise(lit_midpoint = mean(value),
                     lit_min = min(value),
                     lit_max = max(value),
                     lit_source = paste0(unique(sheet),collapse = "; "),
                     area = paste0(unique(area),collapse = "; "),
                     .by = c('set', 'indicator', 'product')) |>
    dplyr::mutate(
      unit = ifelse(grepl("_ha",indicator),"kgCO2e_per_ha","kgCO2e_per_t")
    )  
) |> 
  dplyr::bind_rows(
    lit_GHGE_components |>
      dplyr::mutate(GHGE_kgCO2e_t = GHGE_kgCO2e_kg*10^3) |>
      tidyr::pivot_longer(cols = c(GHGE_kgCO2e_ha,GHGE_kgCO2e_t),
                          names_to = "indicator") |>
      dplyr::filter(!is.na(value)) |>
      dplyr::mutate(indicator = paste(indicator,Harmonized_Category,sep = "__")) |>
      dplyr::summarise(lit_midpoint = mean(value, na.rm = T),
                       lit_min = min(value, na.rm = T),
                       lit_max = max(value, na.rm = T),
                       lit_source = paste0(unique(sheet),collapse = "; "),
                       area = paste0(unique(area),collapse = "; "),
                       .by = c('set', 'indicator', 'product')) |>
      dplyr::mutate(
        unit = ifelse(grepl("_ha",indicator),"kgCO2e_per_ha","kgCO2e_per_t")
      )
    
  ) |>
  # arrange columns
  dplyr::select(set, indicator, unit, product, area, lit_midpoint, lit_min, lit_max, lit_source)

```

```{r ademe_data}

# add global GHGE at farm scale from ADEME data base of carbon diagnosis
ADEME_desc_stat_TF <- read.csv("ADEME_desc_stat_TF.csv")

lit_table <- dplyr::bind_rows(
  lit_table,
  ADEME_desc_stat_TF |>
    dplyr::filter(exploitation_otex %in% c(1500, 4500, 4600)) |>
    dplyr::mutate(
      set = "GHG Emissions",
      indicator = dplyr::case_when(
        variable == "emissions_ges_totales_exploitation" ~ "pseudofarm_total_ghge_kgCO2e",
        variable == "emissions_ges_co2" ~ "pseudofarm_total_CO2_kgCO2e",
        variable == "emissions_ges_ch4" ~ "pseudofarm_total_CH4_kgCO2e",
        variable == "emissions_ges_n2o" ~ "pseudofarm_total_N2O_kgCO2e",
        
        variable == "emissions_ges_totales_exploitation_per_ha" ~ "pseudofarm_total_ghge_kgCO2e_per_ha",
        variable == "emissions_ges_co2_per_ha" ~ "pseudofarm_total_CO2_kgCO2e_per_ha",
        variable == "emissions_ges_ch4_per_ha" ~ "pseudofarm_total_CH4_kgCO2e_per_ha",
        variable == "emissions_ges_n2o_per_ha" ~ "pseudofarm_total_N2O_kgCO2e_per_ha",
        
        variable == "emissions_ges_totales_exploitation_per_LU" ~ "pseudofarm_total_ghge_kgCO2e_per_LU",
        variable == "emissions_ges_co2_per_LU" ~ "pseudofarm_total_CO2_kgCO2e_per_LU",
        variable == "emissions_ges_ch4_per_LU" ~ "pseudofarm_total_CH4_kgCO2e_per_LU",
        variable == "emissions_ges_n2o_per_LU" ~ "pseudofarm_total_N2O_kgCO2e_per_LU"
      ),
      unit = dplyr::case_when(
        grepl("_per_ha$",variable) ~ "kgCO2e_per_ha",
        grepl("_per_LU$",variable) ~ "kgCO2e_per_LU",
        .default = "kgCO2e"
      ), 
      product = dplyr::case_when(
        exploitation_otex == 1500 ~ "Field crop farms",
        exploitation_otex == 4500 ~ "Dairy farms",
        exploitation_otex == 4600 ~ "Beef farms" 
      ),
      area = "France", 
      lit_midpoint = mean,
      lit_min = pmax(mean - sd, 0, na.rm = T),
      lit_max = mean + sd,
      lit_source = "ADEME data base"
    ) |>
    dplyr::filter(!is.na(indicator)) |>
    # arrange columns
    dplyr::select(set, indicator, unit, product, area, lit_midpoint, lit_min, lit_max, lit_source)
)

```


```{r display_tab}
# Display

lit_table |> arrange(set, product, indicator) |> gt()

```


# Package outputs

## Agricultural practices

### Quantity of mineral and organic N per hectare

We retrieve the average quantity of mineral nitrogen per hectare for different crops from Farm Survey Practices, and used it as an allocation key. We can now compare if the average per crops from the package output are close to the retrieved averages from these surveys.

```{r N_ferti}

# select crop farms in France
tmp_object_crop = m_prune_farms(my_object_GHGE, TF8 == 1, COUNTRY == "FRA")

# Retrieve fertilization estimates for crops, filtering to common wheat only
tmp_nferti <- tmp_object_crop@practices$crops$N_ferti |>
  dplyr::filter(FADN_code_letter == "CWHTC")

my_ferti = tibble::tibble(
  indicator = c("mineral fertilization", "organic fertilizer"),
  unit = "kg N / ha",
  product = "wheat",
  estimate = c(mean(tmp_nferti$N_ferti_min, na.rm = T),
               mean(tmp_nferti$N_ferti_org, na.rm = T))
)

```

### Quantity of animal manure per hectare


### DMI, GE and Nex per head of dairy cows

The package estimates the feed Dry Matter Intake (DMI) and Gross Energy (GE) supplied to livestock, and deduce the livestock N excretion rates (Nex).
We can also estimate the Roughage-to-concentrate ratio  (DM basis), %.

```{r feed_param}

# select dairy cattle farms in France
tmp_object_milk = m_prune_farms(my_object_GHGE, TF14 == "45", COUNTRY == "FRA")

# Retrieve model estimates
my_Nexcr <- tmp_object_milk@practices$herding$N_excretion$cattle |>
  # filter for dairy cows
  dplyr::filter(FADN_code_letter == "LCOWDAIR") |>
  # select variables
  dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),
                DM_t_anim,
                GE_DMI_MJ_anim_day,GE_IPCC_MJ_anim_day,
                Nex_DMI_kgN_anim_y,Nex_IPCC_kgN_anim_y) |>
  # add share concent
  dplyr::left_join(
    tmp_object_milk@practices$herding$general |>
      # filter for dairy cows
      dplyr::filter(FADN_code_letter == "LCOWDAIR") |>
      dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),
                    share_concent) |>
      # convert in percent
      dplyr::mutate(share_concent = share_concent*100),
    by = tmp_object_milk@traceability$id_cols
  )

# resume table
my_Nexcr_sumup <- my_Nexcr |>
  dplyr::summarise(dplyr::across(
    c(DM_t_anim,
      GE_DMI_MJ_anim_day,GE_IPCC_MJ_anim_day,
      Nex_DMI_kgN_anim_y,Nex_IPCC_kgN_anim_y,
      share_concent),
    list(
      mean = ~ round(mean(.x, na.rm = T),2),
      sd = ~ round(sd(.x, na.rm = T),2)
    ),
    .names = "{.col}_{.fn}"
  ))

```

### Rearing parameters

For dairy cows, we estimate different rearing parameters such as the age at first calving, the renewal rate, or the number of heifers to renew dairy cows.

```{r rearing_param}

# Herd rearing parameters
herd_rear_param = f_herd_rearing_param_cattle(tmp_object_milk) |>
  dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),
                t_1st_calve,rt_LCOWDAIR,LCOWDAIR_Qobs,LCOWDAIR_Fin) |>
  dplyr::mutate(
    # first_calving_age
    first_calving_age = t_1st_calve,
    # renewal_rate
    ## we convert first calving cow number into livestock unit using the heifer livestock unit coefficient
    renewal_rate = mean(LCOWDAIR_Fin*0.8 / LCOWDAIR_Qobs * 100,na.rm = T)
  )


```


### Quantity of purchased feed/head

source ?

```{r}

```

### Yields

```{r yields}

mean(tmp_object_milk@output$other_herd_products |>
       dplyr::filter(output == "milk" & species == "cattle") |>
       dplyr::pull(yield))



```


## GHGE

### GHGE crops

```{r GHGE_output_wheat}


# Retrieve GHGE estimates for crops, filtering to common wheat only
tmp_ghge_wheat <- tmp_object_crop@footprints$GHGE$GHGE_crops |>
  dplyr::filter(FADN_code_letter == "CWHTC")

(tmp_ghge_wheat)

# Reshape to long format, keeping per-ha and per-t columns, dropping totals
tmp_ghge_wheat_compo_long <- tmp_ghge_wheat |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("per_ha$|per_t$"),      # keep intensity estimates only
    -dplyr::matches("total_ghg")           # drop aggregated totals
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha$|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "N2O")        ~ "Field N2O Emissions",
      str_detect(component, "ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      TRUE                                ~ "Other Inputs & Processes" # pesticides, seeds, etc.
    ),
    product = "wheat",
    # Build composite indicator label encoding unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",  component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on denominator (area vs. product mass)
    unit = dplyr::case_when(
      grepl("per_ha$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",  component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_wheat_compo <- tmp_ghge_wheat_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_crop@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )


```

```{r GHGE_output_maize}


# Retrieve GHGE estimates for crops, filtering to common maize only
tmp_ghge_maize <- tmp_object_crop@footprints$GHGE$GHGE_crops |>
  dplyr::filter(FADN_code_letter == "CMZ")

tmp_ghge_maize

# Reshape to long format, keeping per-ha and per-t columns, dropping totals
tmp_ghge_maize_compo_long <- tmp_ghge_maize |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("per_ha$|per_t$"),      # keep intensity estimates only
    -dplyr::matches("total_ghg")           # drop aggregated totals
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha$|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "N2O")        ~ "Field N2O Emissions",
      str_detect(component, "ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      TRUE                                ~ "Other Inputs & Processes" # pesticides, seeds, etc.
    ),
    product = "maize_grain",
    # Build composite indicator label encoding unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",  component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on denominator (area vs. product mass)
    unit = dplyr::case_when(
      grepl("per_ha$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",  component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_maize_compo <- tmp_ghge_maize_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_crop@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )


```

```{r GHGE_output_rapeseed}


# Retrieve GHGE estimates for crops, filtering to common rapeseed only
tmp_ghge_rapeseed <- tmp_object_crop@footprints$GHGE$GHGE_crops |>
  dplyr::filter(FADN_code_letter == "CRAPE")

tmp_ghge_rapeseed

# Reshape to long format, keeping per-ha and per-t columns, dropping totals
tmp_ghge_rapeseed_compo_long <- tmp_ghge_rapeseed |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("per_ha$|per_t$"),      # keep intensity estimates only
    -dplyr::matches("total_ghg")           # drop aggregated totals
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha$|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "N2O")        ~ "Field N2O Emissions",
      str_detect(component, "ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      TRUE                                ~ "Other Inputs & Processes" # pesticides, seeds, etc.
    ),
    product = "rapeseed",
    # Build composite indicator label encoding unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",  component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on denominator (area vs. product mass)
    unit = dplyr::case_when(
      grepl("per_ha$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",  component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_rapeseed_compo <- tmp_ghge_rapeseed_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_crop@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )

```

### GHGE milk

```{r GHGE_output_milk}

(tmp_object_milk@footprints$GHGE$GHGE_milk)

# Retrieve GHGE estimates for milk, filtering to cow milk output only
tmp_ghge_milk <- tmp_object_milk@footprints$GHGE$GHGE_milk |>
  dplyr::filter(species == "cattle")

# Reshape GHGE milk data to long format, keeping only per-ha (pseudofarm) and per-t columns
tmp_ghge_milk_compo_long <- tmp_ghge_milk |>
  dplyr::select(
    dplyr::all_of(tmp_object_milk@traceability$id_cols),
    dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "milk",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_milk_compo <- tmp_ghge_milk_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_milk@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )


```

### GHGE beef

For beef, as wee compare our values to CAP2ER reports that consider only the farm area, we use farm values, not pseudofarm ones.

```{r GHGE_output_beef}

# select meat farms in France
tmp_object_beef = m_prune_farms(my_object_GHGE, TF14 == "49", COUNTRY == "FRA")

# keep farms with 100% grassland
#quantile(tmp_object_beef@practices$herding$general$ha_perm_grassland_pha)
#tmp_object_beef = m_keep_farms(
#  tmp_object_beef,
#  tmp_object_beef@practices$herding$general |>
#    dplyr::filter(ha_perm_grassland_pha == 1) |>
#    dplyr::select(ID,YEAR,COUNTRY)
#)

#(tmp_object_beef@footprints$GHGE$GHGE_meat)

# Retrieve GHGE estimates for meat, filtering to cow meat output only
tmp_ghge_beef <- tmp_object_beef@footprints$GHGE$GHGE_meat |>
  dplyr::filter(output == "meat_beef")

# Reshape GHGE meat data to long format, keeping only per-ha (farm) and per-t columns
tmp_ghge_beef_compo_long <- tmp_ghge_beef |>
  dplyr::select(
    dplyr::all_of(tmp_object_beef@traceability$id_cols),
    output, species,
    dplyr::matches("per_ha_farm$|per_t$"),   # keep farm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|^feed_pseudofarm") # drop aggregated/herd-level columns
    #dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    #-dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
    
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha_farm|per_t$"),
    #cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = dplyr::case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "cattle_meat",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha_farm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      #grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      grepl("per_ha_farm$", component) ~ "kgCO2e_per_ha",
      #grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_beef_compo <- tmp_ghge_beef_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_beef@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )

tmp_ghge_beef_compo

```

### GHGE pork

```{r GHGE_output_pork}

# select meat farms in France
tmp_object_pork = m_prune_farms(my_object_GHGE, TF14 == "50", COUNTRY == "FRA")

# keep farms with pigs
quantile(tmp_object_pork@herd |> dplyr::filter(FADN_code_letter == "LPIGFAT") |> dplyr::pull(Qobs))
tmp_object_pork = m_keep_farms(
  tmp_object_pork,
  tmp_object_pork@herd |>
    dplyr::filter(FADN_code_letter == "LPIGFAT" & Qobs >0) |>
    dplyr::select(ID,YEAR,COUNTRY)
)

#(tmp_object_pork@footprints$GHGE$GHGE_meat)

# Retrieve GHGE estimates for meat, filtering to cow meat output only
tmp_ghge_pork <- tmp_object_pork@footprints$GHGE$GHGE_meat |>
  dplyr::filter(output == "meat_pork")

summary(tmp_ghge_pork)

# Reshape GHGE meat data to long format, keeping only per-ha (farm) and per-t columns
tmp_ghge_pork_compo_long <- tmp_ghge_pork |>
  dplyr::select(
    dplyr::all_of(tmp_object_pork@traceability$id_cols),
    output, species,
    #dplyr::matches("per_ha_farm$|per_t$"),   # keep farm and per-tonne estimates
    #-dplyr::matches("total_ghg|ghge_herd|^feed_pseudofarm") # drop aggregated/herd-level columns
    dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
    
  ) |>
  tidyr::pivot_longer(
    #cols = dplyr::matches("per_ha_farm|per_t$"),
    cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = dplyr::case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "meat_pork",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      #grepl("per_ha_farm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      #grepl("per_ha_farm$", component) ~ "kgCO2e_per_ha",
      grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_pork_compo <- tmp_ghge_pork_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_pork@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )

tmp_ghge_pork_compo

```

### GHGE chicken

```{r GHGE_output_chicken}

# select meat farms in France
tmp_object_chicken = m_prune_farms(my_object_GHGE, TF14 == "50", COUNTRY == "FRA")

# keep farms with broilers
quantile(tmp_object_chicken@herd |> dplyr::filter(FADN_code_letter == "LPLTRBROYL") |> dplyr::pull(Qobs))
tmp_object_chicken = m_keep_farms(
  tmp_object_chicken,
  tmp_object_chicken@herd |>
    dplyr::filter(FADN_code_letter == "LPLTRBROYL" & Qobs >0) |>
    dplyr::select(ID,YEAR,COUNTRY)
)

#(tmp_object_chicken@footprints$GHGE$GHGE_meat)

# Retrieve GHGE estimates for meat, filtering to cow meat output only
tmp_ghge_chicken <- tmp_object_chicken@footprints$GHGE$GHGE_meat |>
  dplyr::filter(output == "meat_chicken")

summary(tmp_ghge_chicken)

# Reshape GHGE meat data to long format, keeping only per-ha (farm) and per-t columns
tmp_ghge_chicken_compo_long <- tmp_ghge_chicken |>
  dplyr::select(
    dplyr::all_of(tmp_object_chicken@traceability$id_cols),
    output, species,
    #dplyr::matches("per_ha_farm$|per_t$"),   # keep farm and per-tonne estimates
    #-dplyr::matches("total_ghg|ghge_herd|^feed_pseudofarm") # drop aggregated/herd-level columns
    dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
    
  ) |>
  tidyr::pivot_longer(
    #cols = dplyr::matches("per_ha_farm|per_t$"),
    cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = dplyr::case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "meat_chicken",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      #grepl("per_ha_farm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      #grepl("per_ha_farm$", component) ~ "kgCO2e_per_ha",
      grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_chicken_compo <- tmp_ghge_chicken_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_chicken@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )

tmp_ghge_chicken_compo

```

### GHGE eggs

```{r GHGE_output_eggs}

# select meat farms in France
tmp_object_eggs = m_prune_farms(my_object_GHGE, TF14 == "50", COUNTRY == "FRA")

# keep farms that produce eggs
quantile(tmp_object_eggs@output$other_herd_products |> dplyr::filter(output == "eggs") |> dplyr::pull(prod_t))
tmp_object_eggs = m_keep_farms(
  tmp_object_eggs,
  tmp_object_eggs@output$other_herd_products |>
    dplyr::filter(output == "eggs" & prod_t >0) |>
    dplyr::select(ID,YEAR,COUNTRY)
)

#(tmp_object_eggs@footprints$GHGE$GHGE_eggs)

# Retrieve GHGE estimates for meat, filtering to cow meat output only
tmp_ghge_eggs <- tmp_object_eggs@footprints$GHGE$GHGE_eggs |>
  dplyr::filter(output == "eggs")

summary(tmp_ghge_eggs)

# Reshape GHGE meat data to long format, keeping only per-ha (farm) and per-t columns
tmp_ghge_eggs_compo_long <- tmp_ghge_eggs |>
  dplyr::select(
    dplyr::all_of(tmp_object_eggs@traceability$id_cols),
    output, species,
    #dplyr::matches("per_ha_farm$|per_t$"),   # keep farm and per-tonne estimates
    #-dplyr::matches("total_ghg|ghge_herd|^feed_pseudofarm") # drop aggregated/herd-level columns
    dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
    
  ) |>
  tidyr::pivot_longer(
    #cols = dplyr::matches("per_ha_farm|per_t$"),
    cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = dplyr::case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "eggs",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      #grepl("per_ha_farm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      #grepl("per_ha_farm$", component) ~ "kgCO2e_per_ha",
      grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_eggs_compo <- tmp_ghge_eggs_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_eggs@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )

tmp_ghge_eggs_compo

```

### GHGE farm

```{r GHGE_output_farms}

(my_object_GHGE@footprints$GHGE$GHGE_farm)

# Crop farms ----
tmp_ghge_crop_farm <- tmp_object_crop@footprints$GHGE$GHGE_farm |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("^pseudofarm_total")   # keep pseudofarm estimates
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("pseudofarm_total"),
    names_to = "component"
  ) |>
  # add total area
  dplyr::left_join(
    tmp_object_crop@farm |>
      dplyr::select(dplyr::all_of(tmp_object_crop@traceability$id_cols),SE025),
    by = tmp_object_crop@traceability$id_cols
  ) |>
  # divide by total area
  dplyr::mutate(
    value_per_ha = value / dplyr::na_if(SE025,0)
  ) |>
  # estimate averages
  dplyr::summarise(
    estimate = mean(value_per_ha, na.rm = T),
    .by = 'component'
  ) |>
  dplyr::mutate(
    indicator = paste0(component,"_per_ha"),
    unit = "kgCO2e_per_ha", 
    product = "Field crop farms"
  ) |>
  dplyr::select(
    indicator, unit, product, estimate
  )

# Dairy farms ----
tmp_ghge_dairy_farm <- tmp_object_milk@footprints$GHGE$GHGE_farm |>
  dplyr::select(
    dplyr::all_of(tmp_object_milk@traceability$id_cols),
    dplyr::matches("^pseudofarm_total")   # keep pseudofarm estimates
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("pseudofarm_total"),
    names_to = "component"
  ) |>
  # add total LU
  dplyr::left_join(
    tmp_object_milk@herd |>
      dplyr::filter(species == "cattle") |>
      dplyr::summarise(
        tot_LU = sum(ALU, na.rm = T),
        .by =  dplyr::all_of(tmp_object_milk@traceability$id_cols)
      ),
    by = tmp_object_milk@traceability$id_cols
  ) |>
  # divide by total LU
  dplyr::mutate(
    value_per_LU = value / dplyr::na_if(tot_LU,0)
  ) |>
  # estimate averages
  dplyr::summarise(
    estimate = mean(value_per_LU, na.rm = T),
    .by = 'component'
  ) |>
  dplyr::mutate(
    indicator = paste0(component,"_per_LU"),
    unit = "kgCO2e_per_LU", 
    product = "Dairy farms"
  ) |>
  dplyr::select(
    indicator, unit, product, estimate
  )

# Beef farms ----
tmp_ghge_beef_farm <- tmp_object_beef@footprints$GHGE$GHGE_farm |>
  dplyr::select(
    dplyr::all_of(tmp_object_beef@traceability$id_cols),
    dplyr::matches("^pseudofarm_total")   # keep pseudofarm estimates
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("pseudofarm_total"),
    names_to = "component"
  ) |>
  # add total LU
  dplyr::left_join(
    tmp_object_beef@herd |>
      dplyr::filter(species == "cattle") |>
      dplyr::summarise(
        tot_LU = sum(ALU, na.rm = T),
        .by =  dplyr::all_of(tmp_object_beef@traceability$id_cols)
      ),
    by = tmp_object_beef@traceability$id_cols
  ) |>
  
  # divide by total LU
  dplyr::mutate(
    value_per_LU = value / dplyr::na_if(tot_LU,0)
  ) |>
  # estimate averages
  dplyr::summarise(
    estimate = mean(value_per_LU, na.rm = T),
    .by = 'component'
  ) |>
  dplyr::mutate(
    indicator = paste0(component,"_per_LU"),
    unit = "kgCO2e_per_LU", 
    product = "Beef farms"
  ) |>
  dplyr::select(
    indicator, unit, product, estimate
  )



```

## Combine F2F outputs

```{r estimate_table}

# Build agricultural practice indicators for dairy cows from herd nutrition & rearing parameters
F2F_output_agri_pract <- tibble::tribble(
  ~indicator,          ~product,       ~estimate,
  "DMI",               "dairy_cows",   my_Nexcr_sumup$DM_t_anim_mean[[1]],           # dry matter intake (t/animal)
  "GE",                "dairy_cows",   my_Nexcr_sumup$GE_DMI_MJ_anim_day_mean[[1]],  # gross energy intake (MJ/animal/day)
  "Nex",               "dairy_cows",   my_Nexcr_sumup$Nex_DMI_kgN_anim_y_mean[[1]],  # nitrogen excretion (kgN/animal/year)
  "share_concent",     "dairy_cows",   my_Nexcr_sumup$share_concent_mean[[1]],        # concentrate share in diet
  "first_calving_age", "dairy_cows",   mean(herd_rear_param$first_calving_age, na.rm = TRUE) * 12, # convert years to months
  "renewal_rate",      "dairy_cows",   mean(herd_rear_param$renewal_rate, na.rm = TRUE)
) |>
  # Attach units from the literature reference table
  dplyr::left_join(
    lit_agri_pract |> dplyr::select(indicator, unit),
    by = "indicator"
  ) |>
  # add N ferti
  dplyr::bind_rows(
    my_ferti
  )


# Build total GHGE summary (per ha and per tonne)
F2F_output_ghge0 <- dplyr::bind_rows(
  tmp_ghge_wheat |>
    dplyr::select(ID, YEAR, COUNTRY, total_ghg_crop_kgCO2e_per_ha, total_ghg_crop_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = total_ghg_crop_kgCO2e_per_ha,
                  GHGE_kgCO2e_t = total_ghg_crop_kgCO2e_per_t) |>
    dplyr::mutate(product = "wheat"),
  tmp_ghge_maize |>
    dplyr::select(ID, YEAR, COUNTRY, total_ghg_crop_kgCO2e_per_ha, total_ghg_crop_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = total_ghg_crop_kgCO2e_per_ha,
                  GHGE_kgCO2e_t = total_ghg_crop_kgCO2e_per_t) |>
    dplyr::mutate(product = "maize_grain"),
  tmp_ghge_rapeseed |>
    dplyr::select(ID, YEAR, COUNTRY, total_ghg_crop_kgCO2e_per_ha, total_ghg_crop_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = total_ghg_crop_kgCO2e_per_ha,
                  GHGE_kgCO2e_t = total_ghg_crop_kgCO2e_per_t) |>
    dplyr::mutate(product = "rapeseed"),
  tmp_ghge_milk |>
    dplyr::select(ID, YEAR, COUNTRY, pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm, pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm,
                  GHGE_kgCO2e_t = pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::mutate(product = "milk"),
  tmp_ghge_beef |>
    dplyr::select(ID, YEAR, COUNTRY, pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm, pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm,
                  GHGE_kgCO2e_t = pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::mutate(product = "cattle_meat"),
  tmp_ghge_pork |>
    dplyr::select(ID, YEAR, COUNTRY, pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm, pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm,
                  GHGE_kgCO2e_t = pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::mutate(product = "meat_pork"),
  tmp_ghge_chicken |>
    dplyr::select(ID, YEAR, COUNTRY, pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm, pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm,
                  GHGE_kgCO2e_t = pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::mutate(product = "meat_chicken"),
  tmp_ghge_eggs |>
    dplyr::select(ID, YEAR, COUNTRY, pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm, pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::rename(GHGE_kgCO2e_ha = pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm,
                  GHGE_kgCO2e_t = pseudofarm_ghge_herd_kgCO2e_per_t) |>
    dplyr::mutate(product = "eggs")
) |>
  tidyr::pivot_longer(
    cols = c(GHGE_kgCO2e_ha, GHGE_kgCO2e_t),
    names_to = "indicator",
    values_to = "estimate"
  ) |>
  dplyr::mutate(
    unit = ifelse(indicator == "GHGE_kgCO2e_ha", 'kgCO2e_per_ha', 'kgCO2e_per_t')
  )

F2F_output_ghge <- F2F_output_ghge0 |>
  dplyr::filter(is.finite(estimate)) |>
  dplyr::summarise(
    estimate = mean(estimate, na.rm = TRUE),
    .by = c(product, indicator, unit)
  )  |>
  dplyr::bind_rows(tmp_ghge_crop_farm) |>
  dplyr::bind_rows(tmp_ghge_dairy_farm) |>
  dplyr::bind_rows(tmp_ghge_beef_farm)

# Compute mean GHGE by harmonized category (component breakdown)
F2F_output_ghge_compo0 <- dplyr::bind_rows(
  tmp_ghge_wheat_compo,
  tmp_ghge_maize_compo,
  tmp_ghge_rapeseed_compo,
  tmp_ghge_milk_compo,
  tmp_ghge_beef_compo,
  tmp_ghge_pork_compo,
  tmp_ghge_chicken_compo,
  tmp_ghge_eggs_compo
)

## average each GHGE component across farms
F2F_output_ghge_compo <- F2F_output_ghge_compo0 |>
  dplyr::filter(is.finite(value)) |>
  dplyr::summarise(
    estimate = mean(value, na.rm = TRUE),
    .by = c(product, indicator, unit)
  )

# Stack all output tables (practices, total GHGE, GHGE components) into a single flat table
F2F_output <- Reduce(
  function(x, y) dplyr::bind_rows(x, y),
  list(F2F_output_agri_pract,
       F2F_output_ghge,
       F2F_output_ghge_compo)
)

# Render the final output table sorted by product and indicator
F2F_output |> arrange(product, indicator) |> gt()

```

# Comparison

## Total GHGE

```{r plot_data}

library(dplyr)
library(stringr)
library(ggplot2)
library(forcats)

# 1. Prepare data for plotting
tmp_F2F_values0 <- F2F_output_ghge0 %>%
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = gsub("_per_"," / ", unit),
    # Facet by product_type
    product_type = as.factor(ifelse(grepl("wheat|maize|rapeseed", product), "Crops", "Animal products")),
    product_source = paste(product, "FADN France", sep = " - ")
  ) |>
# Compute 99th percentile for visual winsorisation by unit
  mutate(
    p99 = quantile(estimate, 0.99, na.rm = TRUE),
    .by = c(Category, unit_label, product_type, product_source)
  )

tmp_lit_values0 <- lit_table |>
  dplyr::semi_join(
    tmp_F2F_values0 |>
      dplyr::select(product, indicator, unit) |>
      dplyr::distinct(),
    by = c('indicator', 'unit', 'product')
  ) |>
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = gsub("_per_"," / ", unit),
    # Facet by product_type
    product_type = as.factor(ifelse(grepl("wheat|maize|rapeseed", product), "Crops", "Animal products")),
     product_source = paste(product, "Literature", sep = " - ")
  ) 

```

```{r plot_crops}

# 3. Filter product and Winsorisation
tmp_plot_data <- dplyr::bind_rows(
  tmp_F2F_values0 |>
    dplyr::filter(estimate <= p99),
  tmp_lit_values0
) |>
  dplyr::filter(product_type == "Crops")


# 2. Create the plot
ggplot(tmp_plot_data) +
  # accepted range
  geom_errorbar(aes(x = product_source,
                    ymin = pmin(lit_midpoint, lit_min, na.rm = T) / 2, 
                    ymax = pmax(lit_midpoint, lit_max, na.rm = T) * 2
                ), colour = "#d8b365",
                linewidth=1,width = 1,position = position_dodge(width = 0.75)) +
  # literature values
  geom_point(aes(x = product_source, y = lit_midpoint), colour = "#5ab4ac", 
             shape = 1, size=3, stroke = 1, position = position_dodge(width = 0.75)) +
  # min and max
  geom_errorbar(aes(x = product_source, ymin = lit_min, ymax = lit_max), colour = "#5ab4ac",
                linewidth=1,width = 1,position = position_dodge(width = 0.75)) +
  # estimates
  geom_violin(aes(x = product_source, y = estimate, alpha = 0),
              colour = "#999999"
  ) +
  geom_boxplot(aes(x = product_source, y = estimate, alpha = 0),
               colour = "#333333", width = 0.5
  ) +
  # theme
  coord_flip() +
  facet_wrap(vars(unit_label), scales = "free",ncol = 1) +
  labs(y = "GHGE per crop",x= "Products",
       caption = "For visual readability, FADN France estimates are capped at the 99th percentile.") +
  scale_fill_viridis_d() +
  #  scale_fill_manual(values = tmp_FQS_names$colors,labels = tmp_FQS_names$labels,) +
  #  scale_colour_manual(values = tmp_FQS_names$colors,labels = tmp_FQS_names$labels) +
  theme_light() +
  theme(
    legend.position = c(0,-1),legend.direction = "horizontal",legend.title = element_blank(),
    strip.background = element_rect(fill = NA, colour = "#999999"),strip.text = element_text(color = "#333333"))

```

```{r plot_anim_prod}

# 3. Filter product and Winsorisation
tmp_plot_data <- dplyr::bind_rows(
  tmp_F2F_values0 |>
    dplyr::filter(estimate <= p99),
  tmp_lit_values0
) |>
  dplyr::filter(product_type != "Crops")


# 2. Create the plot
ggplot(tmp_plot_data) +
  # accepted range
  geom_errorbar(aes(x = product_source,
                    ymin = pmin(lit_midpoint, lit_min, na.rm = T) / 2, 
                    ymax = pmax(lit_midpoint, lit_max, na.rm = T) * 2
                ), colour = "#d8b365",
                linewidth=1,width = 1,position = position_dodge(width = 0.75)) +
  # literature values
  geom_point(aes(x = product_source, y = lit_midpoint), colour = "#5ab4ac", 
             shape = 1, size=3, stroke = 1, position = position_dodge(width = 0.75)) +
  # min and max
  geom_errorbar(aes(x = product_source, ymin = lit_min, ymax = lit_max), colour = "#5ab4ac",
                linewidth=1,width = 1,position = position_dodge(width = 0.75)) +
  # estimates
  geom_violin(aes(x = product_source, y = estimate, alpha = 0),
              colour = "#999999"
  ) +
  geom_boxplot(aes(x = product_source, y = estimate, alpha = 0),
               colour = "#333333", width = 0.5
  ) +
  # theme
  coord_flip() +
  facet_wrap(vars(unit_label), scales = "free",ncol = 1) +
  labs(y = "GHGE per animal product",x= "Products",
       caption = "For visual readability, FADN France estimates are capped at the 99th percentile.") +
  scale_fill_viridis_d() +
  #  scale_fill_manual(values = tmp_FQS_names$colors,labels = tmp_FQS_names$labels,) +
  #  scale_colour_manual(values = tmp_FQS_names$colors,labels = tmp_FQS_names$labels) +
  theme_light() +
  theme(
    legend.position = c(0,-1),legend.direction = "horizontal",legend.title = element_blank(),
    strip.background = element_rect(fill = NA, colour = "#999999"),strip.text = element_text(color = "#333333"))

```

```{r outliers_per_t}

tmp_outliers = head(tmp_F2F_values0 |>
                      dplyr::filter(grepl("per_t", unit)) |>
                      dplyr::arrange(-estimate),
                    n = 15)

tmp_outliers

tmp_outliers_obj = m_keep_farms(
  my_object_GHGE,
  tmp_outliers
)

# compare feed
dplyr::left_join(
  my_object_GHGE@practices$herding$feed$feed_intake$total |>
    dplyr::summarise(
      DM_t_anim = mean(DM_t_anim, na.rm = TRUE),
      GE_MJ_anim = mean(GE_MJ_anim, na.rm =TRUE),
      .by = FADN_code_letter
    ),
  tmp_outliers_obj@practices$herding$feed$feed_intake$total |>
    dplyr::summarise(
      DM_t_anim_outliers = mean(DM_t_anim, na.rm = TRUE),
      GE_MJ_anim_outliers = mean(GE_MJ_anim, na.rm =TRUE),
      .by = FADN_code_letter
    )
)

# compare GHGE components

dplyr::left_join(
  my_object_GHGE@footprints$GHGE$GHGE_meat |>
    dplyr::summarise(dplyr::across(
      .cols = dplyr::matches("kgCO2e"),
      .fns = ~sum(.x, na.rm = TRUE)
    ),
    .by = c(output, species)) |>
    tidyr::pivot_longer(
      cols = dplyr::matches("kgCO2e")
    ),
  
  tmp_outliers_obj@footprints$GHGE$GHGE_meat |>
    dplyr::summarise(dplyr::across(
      .cols = dplyr::matches("kgCO2e"),
      .fns = ~sum(.x, na.rm = TRUE)
    ),
    .by = c(output, species)) |>
    tidyr::pivot_longer(
      cols = dplyr::matches("kgCO2e"),
      values_to = "value_outlier"
    ),
  by = c('output', 'species', 'name')
) |>
  dplyr::filter(!is.na(value_outlier)) |>
  dplyr::mutate(
    diff_p100 = abs(value_outlier - value) / value
  ) |>
  dplyr::arrange(-diff_p100)



```

```{r cattle_meat_with_low_GHGE}

tmp_data = tmp_F2F_values0 |>
                      dplyr::filter(grepl("per_ha", unit),
                                    grepl("cattle_meat", product)) 

tmp_quantile = quantile(tmp_data$estimate)

tmp_outliers_obj = m_keep_farms(
  my_object_GHGE,
  tmp_data |>
  dplyr::filter(estimate < tmp_quantile['25%'])
)

tmp_OK_farms <- m_keep_farms(
  my_object_GHGE,
  tmp_data |>
  dplyr::filter(estimate > tmp_quantile['75%'])
)


```


## Source of GHGE

```{r plot_impact_compo}

# 0. Summary table
F2F_output_ghge_compo0 |>
  summarise(
    n        = n(),
    mean     = mean(value, na.rm = TRUE),
    sd       = sd(value, na.rm = TRUE),
    min      = min(value, na.rm = TRUE),
    q25      = quantile(value, 0.25, na.rm = TRUE),
    median   = median(value, na.rm = TRUE),
    q75      = quantile(value, 0.75, na.rm = TRUE),
    max      = max(value, na.rm = TRUE),
    n_NA     = sum(is.na(value)),
    .by = c(product, unit, indicator)
  )

# 1. Prepare data for plotting
tmp_F2F_values <- F2F_output_ghge_compo0 %>%
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = gsub("_per_"," / ", unit)
  ) 

tmp_lit_values <- lit_table |>
  dplyr::semi_join(
    tmp_F2F_values |>
      dplyr::select(product, unit, indicator) |>
      dplyr::distinct(),
    by = c('product', 'unit', 'indicator')
  ) |>
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = gsub("_per_"," / ", unit)
  ) 


# 2. Create the plot
ggplot() +
  # accepted range
  geom_errorbar(data = tmp_lit_values,
                aes(x = product,
                    ymin = pmin(lit_midpoint, lit_min, na.rm = T) / 2, 
                    ymax = pmax(lit_midpoint, lit_max, na.rm = T) * 2
                ), colour = "lightpink",
                linewidth=1,width = 1,position = position_dodge(width = 0.75)) +
  # literature values
  geom_point(data = tmp_lit_values,
             aes(x = product, y = lit_midpoint), colour = "plum4", 
             shape = 1, size=3, stroke = 1, position = position_dodge(width = 0.75)) +
  # min and max
  geom_errorbar(data = tmp_lit_values,
                aes(x = product, ymin = lit_min, ymax = lit_max), colour = "plum4",
                linewidth=1,width = 1,position = position_dodge(width = 0.75)) +
  # estimates
  geom_violin(data = tmp_F2F_values,
              aes(x = product, y = value, alpha = 0),
              colour = "#999999"
  ) +
  geom_boxplot(data = tmp_F2F_values,
               aes(x = product, y = value, alpha = 0),
               colour = "#333333", width = 0.25
  ) +
  # theme
  coord_flip() +
  facet_wrap(vars(Category, unit_label), scales = "free", ncol = 2) +
  labs(y = "Total GHGE",x= "Products") +
  scale_fill_viridis_d() +
  #  scale_fill_manual(values = tmp_FQS_names$colors,labels = tmp_FQS_names$labels,) +
  #  scale_colour_manual(values = tmp_FQS_names$colors,labels = tmp_FQS_names$labels) +
  theme_light() +
  theme(
    legend.position = c(0,-1),legend.direction = "horizontal",legend.title = element_blank(),
    strip.background = element_rect(fill = NA, colour = "#999999"),strip.text = element_text(color = "#333333"))



```



```{r plot_share}


# 1. Prepare data for plotting

tmp_plot_share <- dplyr::bind_rows(
  lit_table %>%
  # Filter only for the rows that contain the component breakdown (identified by __)
  filter(str_detect(indicator, "__")
         & unit == "kgCO2e_per_t") %>%
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = "kg CO2e / tonne"
  ) |>
    dplyr::filter(indicator %in% F2F_output_ghge_compo$indicator) |>
    dplyr::mutate(
      source = "Literature values",
      value = lit_midpoint
    ),
  F2F_output_ghge_compo %>%
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = gsub("_per_"," / ", unit),
          source = "F2F estimates",
      value = estimate
    )
) |>
  dplyr::filter(
    unit == "kgCO2e_per_t",
    grepl("wheat|milk|cattle_meat|meat_pork", product)
  )

cat(paste("Literature sources: ",paste0(unique(unlist(str_split(tmp_plot_share$lit_source, "; "))), collapse = "; ")))

# 2. Create the plot
ggplot(tmp_plot_share, aes(x = product, y = value, fill = Category)) +
  # Use stat = "identity" to plot the literal values in the table
  geom_bar(stat = "identity", position = "fill", color = "white", linewidth = 0.1) +
  # Facet by Product (rows) and Unit (columns)
  # scales = "free_y" is crucial because ha and t values are not comparable
  facet_grid(~ source) +
  # Styling
  scale_fill_brewer(palette = "Paired") + # Or use scale_fill_viridis_d()
  theme_minimal() +
  labs(
    title = "GHG Emission Components",
    #subtitle = paste("Literature sources: ",paste0(unique(unlist(str_split(tmp_plot_share$lit_source, "; "))), collapse = "; ")),
    x = "Product",
    y = "Shares of GHGE components (% of kg CO2e / t)",
    fill = "Emission Component"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "right",
    #axis.text.x = element_blank(), # Remove x axis text because it's redundant with facets
    axis.ticks.x = element_blank()
  ) +
  guides(fill = guide_legend(ncol = 1, title.position = "top")) +
  coord_flip()




```

# Validation

We considered a check point validated when the estimates is more than half and less than twice the literature midpoint.

```{r validate}

comparison <- validate_against_literature(
  estimates = F2F_output,
  references = lit_table)

# TODO: check missing share concent
# TODO: check lit value for Bellassen

```

```{r summary}

comparison |>
  count(status) |>
  gt() |>
  tab_header(title = "Validation summary")

```

```{r full_table}

comp_table = comparison |>
  dplyr::filter(!is.na(estimate)) |>
  dplyr::select(indicator, product, unit,
                lit_midpoint, lit_min, lit_max, lit_source,
                estimate, rel_deviation_pct, status) |>
  arrange(product,indicator) |>
  gt() |>
  tab_header(
    title    = "Detailed validation results",
    subtitle = paste("Estimation on France data: ",
                     paste(sort(unique(my_object_GHGE@farm$YEAR)),
                           collapse = ", "))
  ) |>
  fmt_number(columns = c(lit_midpoint, lit_min, lit_max, estimate), decimals = 1) |>
  fmt_number(columns = c(rel_deviation_pct), decimals = 1) |>
  tab_style(
    style     = cell_fill(color = "#fdd"),
    locations = cells_body(rows = status == "\U274C out of range")
  ) |>
  tab_style(
    style     = cell_fill(color = "#dfd"),
    locations = cells_body(rows = status == "\U2705 in range")
  ) |>
  tab_style(
    style     = cell_fill(color = "#ffd"),
    locations = cells_body(rows = grepl("tolerance", status))
  )

comp_table

gtsave(comp_table, filename = paste0("F2F_comp_lit_summary_",Sys.Date(),".html"))
gtsave(comp_table, filename = paste0("F2F_comp_lit_summary_",Sys.Date(),".docx"))

```

```{r fig-deviation}
#| fig-cap: "Relative deviation from literature midpoint by indicator"
library(ggplot2)

comparison |>
  filter(!is.na(estimate)) |>
  ggplot(aes(x = reorder(paste(indicator, product, sep = " — "),
                         rel_deviation_pct),
             y = rel_deviation_pct,
             fill = status)) +
  geom_col() +
  geom_hline(aes(yintercept = 50), linetype = "dashed", colour = "red") +
  scale_fill_discrete(direction = -1) +
  coord_flip() +
  labs(x = NULL, y = "Relative deviation from midpoint (%)") +
  theme_minimal()

```

# Checks

```{r check_UAA}

object = my_object_GHGE

id_cols = object@traceability$id_cols

tmp_tot_UAA = object@farm |>
  dplyr::select(dplyr::all_of(id_cols),
                SE025)

tmp_sum_UAA = object@crop |>
  #dplyr::filter(!grepl("CTOT",FADN_code_letter)) |>
  dplyr::summarise(
    sum_UAA = sum(area_ha, na.rm = TRUE),
    .by = dplyr::all_of(id_cols)
  )

tmp_check = dplyr::left_join(
  tmp_tot_UAA,
  tmp_sum_UAA,
  by = id_cols
)

plot(tmp_check$SE025, tmp_check$sum_UAA)
abline(0,1)

```

```{r under_glass}

load("C:/Users/srhuet/OneDrive/Research/GitHub/FADN2Footprint/data_raw/raw_FADN_FRA_18_21.RData")

object = my_object_GHGE

id_cols = object@traceability$id_cols

tmp_tot_UG = my_FADN_data |>
  dplyr::select(dplyr::all_of(id_cols),
                CTOTUG_TO) |>
  dplyr::mutate(ID = as.character(ID),
                YEAR = as.character(YEAR))

tmp_sum_UG = object@crop |>
  dplyr::filter(grepl("UG$",FADN_code_letter)) |>
  dplyr::summarise(
    sum_UG_TO = sum(TO, na.rm = TRUE),
    sum_UG_SV = sum(SV, na.rm = TRUE),
    .by = dplyr::all_of(id_cols)
  )

tmp_check = dplyr::left_join(
  tmp_tot_UG,
  tmp_sum_UG,
  by = id_cols
)

plot(tmp_check$CTOTUG_TO, tmp_check$sum_UG_SV)
abline(0,1)


```



```{r check_LU}

object = my_object_GHGE

id_cols = object@traceability$id_cols


tmp_farm_LU = object@farm |>
  dplyr::select(dplyr::all_of(id_cols),
                SE080)

tmp_herd_LU = object@herd |>
  dplyr::summarise(
    tot_LU = sum(ALU, na.rm = T),
    .by =  dplyr::all_of(id_cols)
  )

tmp_LU = dplyr::left_join(
  tmp_farm_LU,
  tmp_herd_LU,
  by = id_cols
)

plot(tmp_LU$SE080, tmp_LU$tot_LU)

```
## Meat production

I want to compare our estimation for meat production to EUROSTAT price : https://ec.europa.eu/eurostat/databrowser/view/apri_ap_anouta__custom_21870211/default/table
Les prix pour la France sont bizarre : on est ~360€ pour 100kg et par tête pour les génisses alors qu'en Espagne par exemple, on est à 285€ pour 100 kg (donc 2850€/t) et 1200€ par tête
le rendement dans agribalyse est de 1/22e-4=450 kg/ha vs 7.2/15*10^3 = 480 kg/ha dans CAP2ER
```{r meat_price}

# Heiffers

tmp_heif = tmp_object_beef@output$meat |>
  dplyr::filter(FADN_code_letter == "LHEIFFAT")

cat("Average weight for heiffers: ",
    mean(tmp_heif$prod_t / tmp_heif$sales_nb, na.rm = TRUE)," t/head")
cat("\n")
cat("Mean price for heiffer: ",
    mean(tmp_heif$sales_e / tmp_heif$sales_nb, na.rm = TRUE)," €/head and ",
    mean(tmp_heif$sales_e / tmp_heif$prod_t, na.rm = TRUE)," €/t")
cat("\n")
cat("\n")

cat("\nOn average, beef farms sell ",
    (mean(tmp_object_beef@herd |>
            #dplyr::filter(grepl("LHEIF",FADN_code_letter)) |>
            dplyr::pull(share_SSN),
          na.rm = T))*100,"% of their animals for slaughter")


```

I want to compare our liveweight production per livestock units to those reported in CAP2ER => OK, eventhough we overestimate veal live weight

```{r live_weight_production}

object = tmp_object_beef

tmp_lwp <- object@output$meat |>
  dplyr::filter(grepl("meat", output) & species == "cattle") |>
  dplyr::summarise(
    LWP_t = sum(prod_t, na.rm = TRUE),
    .by = dplyr::all_of(id_cols)
  )

tmp_sold_LU = object@herd |>
  dplyr::filter(species == "cattle") |>
  dplyr::summarise(
    tot_LU = sum(SN * livestock_unit_coef, na.rm = T),
    .by =  dplyr::all_of(id_cols)
  )

tmp_lwp_LU <- dplyr::left_join(
  tmp_lwp,
  tmp_sold_LU,
  by = id_cols
) |>
  dplyr::mutate(
    lwp_t_LU = LWP_t / tot_LU
  )

hist(tmp_lwp_LU$lwp_t_LU, main = "Liweweight Production (t/LU)")
quantile(tmp_lwp_LU$lwp_t_LU)
mean(tmp_lwp_LU$lwp_t_LU)

```
```{r}

object@herd |> filter(ID == "52100000050798")

object@output$meat |> filter(ID == "52100000050798")

```


```{r}

my_FADN_data |> filter(ID == "52100000050798") |> dplyr::select(dplyr::all_of(id_cols),dplyr::matches("LBOV")) |> tidyr::pivot_longer(cols = -dplyr::all_of(id_cols))

```

## Eggs

```{r}

# LHENSLAY

tmp_LU <- tmp_object_eggs@herd |>
  dplyr::filter(FADN_code_letter == "LHENSLAY")


tmp_prod <- tmp_object_eggs@output$other_herd_products |>
  dplyr::filter(grepl("PEGG", FADN_code_letter_output)) |>
  dplyr::summarise(
    prod_t = sum(prod_t, na.rm = T),
    .by = c(dplyr::all_of(id_cols), output)
  )

tmp_yield <- dplyr::left_join(
  tmp_LU, tmp_prod,
  by = id_cols
) |>
  dplyr::mutate(
    yield = prod_t / Qobs
  )

quantile(tmp_yield$yield)

```

## Feed

```{r check_purchased_feed_alloc}

object = my_object_GHGE
id_cols = my_object_GHGE@traceability$id_cols

th_feed <- f_feed_theo_ration(object)

# create the "would-be" allocation exactly like f_feed_offfarm does (uses traceability id cols)
feed_purchased_qty <- object@input |>
  # Retrieve farm purchases values for concentrate and rough feed
  dplyr::select(dplyr::all_of(id_cols),
                dplyr::matches("feed_purch")) |>
  ## Purchased concentrated feedstuffs for grazing stock (equines, ruminants) Value in EUR
  ##feed_purch_concent_graz = IGRFEDCNCTRPUR_V,
  ## Purchased coarse fodder for grazing stock (equines, ruminants) Value in EUR
  ##feed_purch_rough_graz = IGRFEDCRSPUR_V,
  ## Purchased feedstuffs for pigs Value in EUR
  ##feed_purch_concent_pigs = IPIGFEDPUR_V,
  ## Purchased feedstuffs for poultry and other small animals Value in EUR
  ##feed_purch_concent_poultry = IPLTRFEDPUR_V,
  tidyr::pivot_longer(cols = tidyselect::matches("feed_purch"),
                      names_to = "feed_var",
                      values_to = "feed_purch_value") |>
  dplyr::filter(feed_purch_value >0) |>
  dplyr::mutate(
    feed_type = dplyr::case_when(
      feed_var == "feed_purch_rough_graz" ~ "feed_rough",
      .default = "feed_concent"
    ),
    species = dplyr::case_when(
      feed_var %in% c("feed_purch_concent_graz","feed_purch_rough_graz") ~ "grazing",
      feed_var == "feed_purch_concent_pigs" ~ "swine",
      feed_var == "feed_purch_concent_poultry" ~ "poultry"
    )
  ) |>
  #filter(feed_purch_value >0) |> # keep null purchases for traceability
  # Add theoretical values
  dplyr::left_join(th_feed$th_feed_value |>
                     dplyr::select(dplyr::all_of(id_cols),
                                   species, FADN_code_letter,
                                   feed_type, Sailley_feed,
                                   th_crop_value_p100_feed_type, euros_t),
                   by = c(id_cols,"feed_type","species")) |>
  # Estimate share of purchased feed value per livestock category and per crop
  dplyr::mutate(
    feed_purch_value_share = feed_purch_value * th_crop_value_p100_feed_type
  ) |>
  # Convert estimated value in ton
  dplyr::mutate(
    feed_purch_t_livcat = feed_purch_value_share / euros_t
  )

# summary per feed_var
tmp_comp <- feed_purchased_qty |>
  dplyr::summarise(
    total_input = mean(feed_purch_value, na.rm=TRUE),
    total_allocated = sum(feed_purch_value_share, na.rm=TRUE),
    n_rows = n(),
    #n_unallocated = sum(is.na(feed_purch_value_share)),
    #.by = feed_var
    .by = c(dplyr::all_of(id_cols), species, feed_type)
  ) |>
  dplyr::mutate(
    diff_p100 = abs(total_allocated - total_input) / total_input,
    diff = ifelse(round(diff_p100,5) > 0, TRUE, FALSE))

table(tmp_comp$diff)

# see if livestock corresponding to feed
tmp_comp |> 
  dplyr::filter(diff) |>
  dplyr::left_join(
    object@herd |> 
      #dplyr::mutate(
      #    species = case_when(
      #        species %in% c("cattle","sheep","goats","horse") ~ "grazing",
      #        .default = species
      #    )
      #) |>
      dplyr::summarise(
        Qobs = sum(Qobs),
        .by = c(dplyr::all_of(id_cols), species)
      ) |>
      tidyr::pivot_wider(names_from = species, values_from = Qobs),
    #by = c(id_cols, 'species')
    by = c(id_cols)
  ) #|> View()




```

Some farms purchase poultry feed but do not have poultry (some do not have livestock at all sometimes)

## Heating fuels

```{r}

semi_join(
  my_object_GHGE@footprints$GHGE$GHGE_herd |>
    dplyr::summarise(ghg_heat_fuel_kgCO2e_livcat = sum(ghg_heat_fuel_kgCO2e_livcat, na.rm = T),
                     .by = dplyr::all_of(id_cols)) |>
    dplyr::filter(ghg_heat_fuel_kgCO2e_livcat == 0),
  my_object_GHGE@input |> dplyr::filter(heating_fuels_l >0)
) #|> View()

tmp_farms = semi_join(
  my_object_GHGE@footprints$GHGE$GHGE_herd |>
    dplyr::summarise(ghg_heat_fuel_kgCO2e_livcat = sum(ghg_heat_fuel_kgCO2e_livcat, na.rm = T),
                     .by = dplyr::all_of(id_cols)) |>
    dplyr::filter(ghg_heat_fuel_kgCO2e_livcat == 0),
  my_object_GHGE@input |> dplyr::filter(heating_fuels_l >0)
) |>
  dplyr::select(dplyr::all_of(my_object_GHGE@traceability$id_cols))

tmp_obbj = m_keep_farms(my_object_GHGE, tmp_farms)

tmp_econ_alloc = f_output_econ_alloc(tmp_obbj)

summary(tmp_econ_alloc$all_outputs)

tmp_econ_alloc$all_outputs #|> View()

```

## Farms with zero GHGE

```{r check_GHGE_farms}

object = my_object_GHGE
id_cols = object@traceability$id_cols

tmp_obj_farms_zero <- m_keep_farms(
  object,
  object@footprints$GHGE$GHGE_farm |>
    dplyr::filter(is.na(farm_total_ghge_kgCO2e ) | farm_total_ghge_kgCO2e  == 0)
)

table(tmp_obj_farms_zero@farm$TF14)

tmp_obj_farms_zero@herd |>
  ggplot() +
  aes(x = FADN_code_letter,
      y = SN) +
  geom_boxplot()

my_FADN_data |>
  dplyr::select(dplyr::all_of(id_cols),
                dplyr::matches("PCNT")) |>
  dplyr::mutate(ID = as.character(ID),
                YEAR = as.character(YEAR)) |>
  dplyr::semi_join(tmp_obj_farms_zero@farm,
                   by = id_cols) |>
  tidyr::pivot_longer(cols = -dplyr::all_of(id_cols)) |>
  dplyr::filter(grepl("_SV$", name))|>
  ggplot() +
  aes(x = name,
      y = value) +
  geom_boxplot()

tmp_obj_farms_zero@herd |>
  ggplot() +
  aes(x = FADN_code_letter,
      y = Qobs) +
  geom_boxplot()

```

## Regressions

```{r}



```

