---
title: "FADN2Footprint — Validation Report"
subtitle: "Comparison of package estimates with national inventoriess"
author: "Sarah HUET & Valentin BELLASSEN"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true        
  pdf:
    toc: true
execute:
  echo: true
  warning: false
editor: 
  markdown: 
    wrap: 150
bibliography: ../../vignettes/references.bib
link-citations: true
---

# Purpose

This report compares the outputs of `FADN2Footprint` (v`r packageVersion("FADN2Footprint")`)
against the national inventories.
An estimate is flagged **in range** if it falls within the literature min–max,
**within tolerance** if the relative deviation from the literature midpoint is
below a per-indicator threshold, or **out of range** otherwise.

# Setup

```{r setup}

library(dplyr)
library(tidyr)
library(stringr)
library(gt)

library(readxl)
library(zip)

library(ggplot2)

devtools::load_all()
#library(FADN2Footprint)

```

# UNFCCC data

## Extract raw data

```{r UNFCCC_raw_data, echo=FALSE, message=FALSE}

# Raw UNFCCC files
folder_path = "C:/Users/srhuet/OneDrive/Research/Data/UNFCCC_ghg_inventories/"


# Wrangle UNFCCC files ----
UNFCCC_data_validation <- list(
     "Summary1.As2" = tibble(),
     "Table3" = tibble(),
     "Table3.D" = tibble()
)

#for (loop_zip in list.files(folder_path)) {
loop_zip = list.files(folder_path)[12]


loop_files = zip_list(paste0(folder_path,loop_zip))
loop_data_country = toupper(str_sub(loop_zip,1,3))
print(loop_data_country)

for (loop_data_year in c("2018", "2019", "2020")) {
     #loop_data_year = sub(paste0("^",loop_data_country,"_2023_(\\d{4})_.*$"), "\\1", loop_files$filename)[29]
     print(loop_data_year)
     
     
     loop_unzip <- tempfile()
     loop_unzip <- utils::unzip(zipfile =  paste0(folder_path,loop_zip),
                                files = loop_files$filename[grep(paste0("_",loop_data_year,"_"),loop_files$filename)])
     
     ## SUMMARY 1.A  SUMMARY REPORT FOR NATIONAL GREENHOUSE GAS INVENTORIES ----
     #sheet Summary1.As2
     
     # read file
     loop_raw_sheet <- read_excel(loop_unzip,sheet = "Summary1.As2", col_names = F)
     # select table
     loop_range <- loop_raw_sheet[seq(which(grepl("Agriculture",loop_raw_sheet$...1)),max(which(is.na(loop_raw_sheet$...1)))-1),]
     # retrieve column names
     loop_colnames <- as.character(loop_raw_sheet[which(grepl("SOURCE",loop_raw_sheet$...1)),])
     colnames(loop_range) <- loop_colnames
     
     # frame tibble
     loop_table <- as_tibble(loop_range) |>
          # numeric values
          dplyr::mutate(dplyr::across(
               .cols = dplyr::all_of(loop_colnames[-1]),
               .fns = ~ suppressWarnings(as.numeric(.))
          )) |>
          # add country and year
          dplyr::mutate(
               Country_ISO_3166_1_A3 = loop_data_country,
               YEAR = as.character(loop_data_year)
          )
     
     UNFCCC_data_validation$Summary1.As2 <- dplyr::bind_rows(UNFCCC_data_validation$Summary1.As2, loop_table)
     
     ## TABLE 3 SECTORAL REPORT FOR AGRICULTURE ----
     #sheet Table3s1 Table3s2
     
     # read file
     loop_raw_sheet <- read_excel(loop_unzip,sheet = "Table3s1", col_names = F)
     # select table
     loop_range <- loop_raw_sheet[seq(which(grepl("Total agriculture",loop_raw_sheet$...1)),max(which(is.na(loop_raw_sheet$...1)))-1),]
     # retrieve column names
     loop_colnames <- as.character(loop_raw_sheet[which(grepl("SOURCE",loop_raw_sheet$...1)),])
     colnames(loop_range) <- loop_colnames
     
     # frame tibble
     loop_table <- as_tibble(loop_range) |>
          # numeric values
          dplyr::mutate(dplyr::across(
               .cols = dplyr::all_of(loop_colnames[-1]),
               .fns = ~ suppressWarnings(as.numeric(.))
          )) |>
          # add country and year
          dplyr::mutate(
               Country_ISO_3166_1_A3 = loop_data_country,
               YEAR = as.character(loop_data_year)
          )
     
     UNFCCC_data_validation$Table3 <- dplyr::bind_rows(UNFCCC_data_validation$Table3, loop_table)
     
     # read file
     loop_raw_sheet <- read_excel(loop_unzip,sheet = "Table3s2", col_names = F)
     # select table
     loop_range <- loop_raw_sheet[seq(which(grepl("Rice cultivation",loop_raw_sheet$...1)),which(grepl("J.  Other",loop_raw_sheet$...1))),]
     # retrieve column names
     loop_colnames <- as.character(loop_raw_sheet[which(grepl("SOURCE",loop_raw_sheet$...1)),])
     colnames(loop_range) <- loop_colnames
     
     # frame tibble
     loop_table <- as_tibble(loop_range) |>
          # numeric values
          dplyr::mutate(dplyr::across(
               .cols = dplyr::all_of(loop_colnames[-1]),
               .fns = ~ suppressWarnings(as.numeric(.))
          )) |>
          # add country and year
          dplyr::mutate(
               Country_ISO_3166_1_A3 = loop_data_country,
               YEAR = as.character(loop_data_year)
          )
     
     UNFCCC_data_validation$Table3 <- dplyr::bind_rows(UNFCCC_data_validation$Table3, loop_table)
     
     ## TABLE 3.D Direct and indirect N2O emissions from agricultural soils ----
     # sheet Table3.D
     
     # read file
     loop_raw_sheet <- read_excel(loop_unzip,sheet = "Table3.D", col_names = F)
     # select table
     loop_range <- loop_raw_sheet[seq(which(grepl("Direct N2O emissions",loop_raw_sheet$...1)),
                                      which(grepl("Nitrogen leaching and run-off",loop_raw_sheet$...1))),]
     # retrieve column names
     loop_colnames <- t(loop_raw_sheet[c( which(grepl("GREENHOUSE GAS",loop_raw_sheet$...1)):
                                               (which(grepl("GREENHOUSE GAS",loop_raw_sheet$...1))+2)),]) |>
          transform(V0 = paste(coalesce(V1,""),coalesce(V2,""),coalesce(V3,"")))
     colnames(loop_range) <- loop_colnames$V0
     
     # frame tibble
     loop_table <- as_tibble(loop_range) |>
          # numeric values
          dplyr::mutate(dplyr::across(
               .cols = dplyr::all_of(loop_colnames$V0[-c(1,2)]),
               .fns = ~ suppressWarnings(as.numeric(.))
          )) |>
          # add country and year
          dplyr::mutate(
               Country_ISO_3166_1_A3 = loop_data_country,
               YEAR = as.character(loop_data_year)
          )
     
     UNFCCC_data_validation$Table3.D <- dplyr::bind_rows(UNFCCC_data_validation$Table3.D, loop_table)
     
     
     
     unlink(loop_unzip)
}
#}

#
#
# TABLE 3 SECTORAL BACKGROUND DATA FOR AGRICULTURE
# 3.A Enteric Fermentation 
# sheet Table3.As1
#
# 3.B(a) CH4 Emissions from Manure Management
# sheet Table3.B(a)s1
#
# 3.B(b) N2O Emissions from Manure Management
# sheet Table3.B(b)
#

rm(list = names(.GlobalEnv)[grepl("loop_", names(.GlobalEnv))])
rm(folder_path)

```

# Package output

## Run package

```{r load_data, echo=FALSE, message=TRUE, warning=FALSE}

## Load raw data and run model
#source("validation_FADN_data.R")
#source("validation_FADN_data_offline.R")
## 
#load("../../data_raw/FADN_FRA_2018_2021_obj_practices.RData")
load("../../data_raw/FADN_FRA_2018_2021_obj_GHGE.RData")
#my_object_GHGE <- m_prune_farms(my_object_GHGE,
#                                YEAR %in% c("2018", "2019", "2020"))

id_cols = my_object_GHGE@traceability$id_cols

cat("FADN dataset loaded:", nrow(my_object_GHGE@farm), "farm-year observations\n")
cat("Countries covered:", paste(unique(my_object_GHGE@farm$COUNTRY), collapse = ", "), "\n")
cat("Years covered:", paste(sort(unique(my_object_GHGE@farm$YEAR)), collapse = ", "), "\n")

```
## Agregate outputs

```{r crops}

# a. Direct N2O emissions from managed soils 	
# 1.   Inorganic N fertilizers(3)	N input from application of inorganic fertilizers to cropland and grassland
crops_n_min_input <- my_object_GHGE@practices$crops$N_ferti |>
     # add agregation weight
     dplyr::left_join(my_object_GHGE@farm |>
                           dplyr::select(dplyr::all_of(id_cols), SYS02),
                      by = id_cols) |>
     # sum up by country and year
     dplyr::summarise(
          N_ferti_min = sum(N_ferti_min * area_ha * SYS02, na.rm = TRUE),
          .by = c(COUNTRY, YEAR)
     )


# 2.   Organic N fertilizers(3)	N input from organic N fertilizers to cropland and grassland
#       a. Animal manure applied to soils	N input from manure applied to soils
#       b. Sewage sludge applied to soils	N input from sewage sludge applied to soils
#       c. Other organic fertilizers applied to soils	N input from application of other organic fertilizers
# 3.   Urine and dung deposited by grazing animals	N excretion on pasture, range and paddock
# 4.   Crop residues	N in crop residues returned to soils
# 5.  Mineralization/immobilization associated with loss/gain of soil organic matter (4)(5)	N in mineral soils that is mineralized in association with loss # of soil C
# 6.   Cultivation of organic soils (i.e. histosols)(2)	Area of cultivated organic soils
# 7.   Other	


# b. Indirect N2O Emissions from managed soils 	
# 1.   Atmospheric deposition(6)	Volatilized N from agricultural inputs of N
# 2.   Nitrogen leaching and run-off	N from fertilizers and other agricultural inputs that is lost through leaching and run-off



```



