---
title: "FADN2Footprint — Validation Report"
subtitle: "Comparison of package estimates with literature values"
author: "Your name"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true        
  pdf:
    toc: true
execute:
  echo: true
  warning: false
editor: 
  markdown: 
    wrap: 150
bibliography: ../../vignettes/references.bib
link-citations: true
---

# Purpose

This report compares the outputs of `FADN2Footprint` (v`r packageVersion("FADN2Footprint")`)
against published literature ranges.
An estimate is flagged **in range** if it falls within the literature min–max,
**within tolerance** if the relative deviation from the literature midpoint is
below a per-indicator threshold, or **out of range** otherwise.

# Setup

```{r setup}

library(dplyr)
library(gt)

#devtools::document()
#devtools::load_all()
library(FADN2Footprint)

```

# Produce package estimates

```{r load_data, echo=FALSE, message=TRUE, warning=FALSE}

## Load raw data and run model
source("validation_FADN_data.R")
#source("validation_FADN_data_offline.R")
## 
#load("../../data_raw/FADN_FRA_2018_2021_obj_GHGE.RData")

cat("FADN dataset loaded:", nrow(my_object_GHGE@farm), "farm-year observations\n")
cat("Countries covered:", paste(unique(my_object_GHGE@farm$COUNTRY), collapse = ", "), "\n")
cat("Years covered:", paste(sort(unique(my_object_GHGE@farm$YEAR)), collapse = ", "), "\n")

```

# Literature references

```{r load_literature, message=FALSE}

lit_GHGE_total <- readxl::read_excel(
  "validation_references.xlsx",
  sheet = "summary_total_GHGE"
) |>
  # filter for wheat and milk in France
  dplyr::filter(area %in% c("France","Europe") #data_extra$country_names$country_name
                & product %in% c("wheat","milk")) |>
  # filter for references with both per ha and per t
  dplyr::filter(GHGE_kgCO2e_kg >0 
                | GHGE_kgCO2e_ha >0) |>
  # add set
  dplyr::mutate(set = "GHG Emissions")

lit_GHGE_components <- readxl::read_excel(
  "validation_references.xlsx",
  sheet = "summary_share_components"
) |>
  # filter for wheat and milk in France
  dplyr::filter(area %in% c("France","Europe") #data_extra$country_names$country_name
                & product %in% c("wheat","milk")) |>
  # remove unclassified components
  dplyr::filter(!is.na(Harmonized_Category)) |>
  # sum up by product,Harmonized_Category,sheet,area
  dplyr::summarise(
    dplyr::across(
      c(GHGE_kgCO2e_ha,GHGE_kgCO2e_kg),
      ~ sum(.x)
    ),
    .by = c('product','Harmonized_Category','sheet','area')
  ) |>
  # add set
  dplyr::mutate(set = "GHG Emissions")

cat("Literature (total GHGE) loaded:", nrow(lit_GHGE_total), "rows\n")
cat("Literature (components of GHGE) loaded:", nrow(lit_GHGE_components), "rows\n")

lit_agri_pract <- readxl::read_excel(
  "validation_references.xlsx",
  sheet = "summary_agri_practices"
)

# add EPK averages for wheat fertilization
tmp = data_extra$PKGC_N_ferti |>
  dplyr::filter(EPK_ferti_crop_name == "Ble_tendre" & ORGANIC == 0) |>
  dplyr::select(N_min,Nmin_sd_w,N_org,Norg_sd_w)

lit_agri_pract <- lit_agri_pract |>
  dplyr::bind_rows(
    tibble::tibble(
      indicator = c("mineral fertilization", "organic fertilizer"),
      unit = "kg N / ha",
      product = "wheat",
      area = "France",
      lit_midpoint = c(tmp$N_min, tmp$N_org),
      lit_min = 0,
      lit_max = c((tmp$N_min + tmp$Nmin_sd_w), (tmp$N_org + tmp$Norg_sd_w)),
      #tolerance_pct = "More than half\n&\nLess than twice",
      lit_source = "EPK 2017"
    )
  ) |>
  # add set
  dplyr::mutate(set = "Agricultural practices")


cat("Literature (agricultural practices) loaded:", nrow(lit_agri_pract), "rows\n")

lit_GHGE_total |> arrange(set,product,sheet) |> gt()
lit_GHGE_components |> arrange(set,product,Harmonized_Category,sheet) |> gt()
lit_agri_pract |> arrange(set,indicator) |> gt()

```


```{r lit_table, message=FALSE}

lit_table = dplyr::bind_rows(
  lit_agri_pract,
  lit_GHGE_total |>
    dplyr::mutate(GHGE_kgCO2e_t = GHGE_kgCO2e_kg*10^3) |>
    tidyr::pivot_longer(cols = c(GHGE_kgCO2e_ha,GHGE_kgCO2e_t),
                        names_to = "indicator") |>
    dplyr::filter(!is.na(value)) |>
    dplyr::summarise(lit_midpoint = mean(value),
                     lit_min = min(value),
                     lit_max = max(value),
                     lit_source = paste0(unique(sheet),collapse = "; "),
                     area = paste0(unique(area),collapse = "; "),
                     .by = c('set', 'indicator', 'product')) |>
    dplyr::mutate(
      unit = ifelse(grepl("_ha",indicator),"kgCO2e_per_ha","kgCO2e_per_t")
    )  
) |> 
  dplyr::bind_rows(
    lit_GHGE_components |>
      dplyr::mutate(GHGE_kgCO2e_t = GHGE_kgCO2e_kg*10^3) |>
      tidyr::pivot_longer(cols = c(GHGE_kgCO2e_ha,GHGE_kgCO2e_t),
                          names_to = "indicator") |>
      dplyr::filter(!is.na(value)) |>
      dplyr::mutate(indicator = paste(indicator,Harmonized_Category,sep = "__")) |>
      dplyr::summarise(lit_midpoint = mean(value, na.rm = T),
                       lit_min = min(value, na.rm = T),
                       lit_max = max(value, na.rm = T),
                       lit_source = paste0(unique(sheet),collapse = "; "),
                       area = paste0(unique(area),collapse = "; "),
                       .by = c('set', 'indicator', 'product')) |>
      dplyr::mutate(
        unit = ifelse(grepl("_ha",indicator),"kgCO2e_per_ha","kgCO2e_per_t")
      )
    
  ) |>
  # arrange columns
  dplyr::select(set, indicator, unit, product, area, lit_midpoint, lit_min, lit_max, lit_source)

```

```{r ademe_data}

# add global GHGE at farm scale from ADEME data base of carbon diagnosis
ADEME_desc_stat_TF <- read.csv("ADEME_desc_stat_TF.csv")

lit_table <- dplyr::bind_rows(
  lit_table,
  ADEME_desc_stat_TF |>
    dplyr::filter(exploitation_otex %in% c(1500,4500)) |>
    dplyr::mutate(
      set = "GHG Emissions",
      indicator = dplyr::case_when(
        variable == "emissions_ges_totales_exploitation" ~ "pseudofarm_total_ghge_kgCO2e",
        variable == "emissions_ges_co2" ~ "pseudofarm_total_CO2_kgCO2e",
        variable == "emissions_ges_ch4" ~ "pseudofarm_total_CH4_kgCO2e",
        variable == "emissions_ges_n2o" ~ "pseudofarm_total_N2O_kgCO2e",
        
        variable == "emissions_ges_totales_exploitation_per_ha" ~ "pseudofarm_total_ghge_kgCO2e_per_ha",
        variable == "emissions_ges_co2_per_ha" ~ "pseudofarm_total_CO2_kgCO2e_per_ha",
        variable == "emissions_ges_ch4_per_ha" ~ "pseudofarm_total_CH4_kgCO2e_per_ha",
        variable == "emissions_ges_n2o_per_ha" ~ "pseudofarm_total_N2O_kgCO2e_per_ha",
        
        variable == "emissions_ges_totales_exploitation_per_LU" ~ "pseudofarm_total_ghge_kgCO2e_per_LU",
        variable == "emissions_ges_co2_per_LU" ~ "pseudofarm_total_CO2_kgCO2e_per_LU",
        variable == "emissions_ges_ch4_per_LU" ~ "pseudofarm_total_CH4_kgCO2e_per_LU",
        variable == "emissions_ges_n2o_per_LU" ~ "pseudofarm_total_N2O_kgCO2e_per_LU"
      ),
      unit = dplyr::case_when(
        grepl("_per_ha$",variable) ~ "kgCO2e_per_ha",
        grepl("_per_LU$",variable) ~ "kgCO2e_per_LU",
        .default = "kgCO2e"
      ), 
      product = ifelse(exploitation_otex == 1500, "Field crop farms", "Dairy farms"),
      area = "France", 
      lit_midpoint = mean,
      lit_min = pmax(mean - sd, 0, na.rm = T),
      lit_max = mean + sd,
      lit_source = "ADEME data base"
    ) |>
    dplyr::filter(!is.na(indicator)) |>
    # arrange columns
    dplyr::select(set, indicator, unit, product, area, lit_midpoint, lit_min, lit_max, lit_source)
)

```


```{r display_tab}
# Display

lit_table |> arrange(set, product, indicator) |> gt()

```


```{r plot_lit_values, message=FALSE}

library(ggplot2)

# 1. Prepare data for plotting
plot_data <- lit_table %>%
  # Filter only for the rows that contain the component breakdown (identified by __)
  filter(str_detect(indicator, "__")
         & unit == "kgCO2e_per_t") %>%
  # Extract the clean category name from the indicator string
  mutate(
    Category = str_extract(indicator, "(?<=__).*"),
    # Make unit names prettier for the plot labels
    unit_label = "kg CO2e / tonne"
  )

# 2. Create the plot
ggplot(plot_data, aes(x = product, y = lit_midpoint, fill = Category)) +
  # Use stat = "identity" to plot the literal values in the table
  geom_bar(stat = "identity", position = "fill", color = "white", linewidth = 0.1) +
  # Facet by Product (rows) and Unit (columns)
  # scales = "free_y" is crucial because ha and t values are not comparable
  facet_grid(~ unit_label) +
  # Styling
  scale_fill_brewer(palette = "Paired") + # Or use scale_fill_viridis_d()
  theme_minimal() +
  labs(
    title = "Literature GHG Emission Sources",
    subtitle = paste("Sources: ",paste0(unique(unlist(str_split(plot_data$lit_source, "; "))), collapse = "; ")),
    x = "Product",
    y = "GHG Emissions (kg CO2e)",
    fill = "Emission Component"
  ) +
  theme(
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "right",
    #axis.text.x = element_blank(), # Remove x axis text because it's redundant with facets
    axis.ticks.x = element_blank()
  ) +
  guides(fill = guide_legend(ncol = 2, title.position = "top"))


```


# Package outputs

## Agricultural practices

### Quantity of mineral and organic N per hectare

We retrieve the average quantity of mineral nitrogen per hectare for different crops from Farm Survey Practices, and used it as an allocation key. We can now compare if the average per crops from the package output are close to the retrieved averages from these surveys.

```{r N_ferti}

# select crop farms in France
tmp_object_crop = m_prune_farms(my_object_GHGE, TF8 == 1, COUNTRY == "FRA")

# Retrieve fertilization estimates for crops, filtering to common wheat only
tmp_nferti <- tmp_object_crop@practices$crops$N_ferti |>
  dplyr::filter(FADN_code_letter == "CWHTC")

my_ferti = tibble::tibble(
  indicator = c("mineral fertilization", "organic fertilizer"),
  unit = "kg N / ha",
  product = "wheat",
  estimate = c(mean(tmp_nferti$N_ferti_min, na.rm = T),
               mean(tmp_nferti$N_ferti_org, na.rm = T))
)

```

### Quantity of animal manure per hectare


### DMI, GE and Nex per head of dairy cows

The package estimates the feed Dry Matter Intake (DMI) and Gross Energy (GE) supplied to livestock, and deduce the livestock N excretion rates (Nex).
We can also estimate the Roughage-to-concentrate ratio  (DM basis), %.

```{r feed_param}

# select dairy cattle farms in France
tmp_object_milk = m_prune_farms(my_object_GHGE, TF14 == "45", COUNTRY == "FRA")

# Retrieve model estimates
my_Nexcr <- tmp_object_milk@practices$herding$N_excretion$cattle |>
  # filter for dairy cows
  dplyr::filter(FADN_code_letter == "LCOWDAIR") |>
  # select variables
  dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),
                DM_t_anim,
                GE_DMI_MJ_anim_day,GE_IPCC_MJ_anim_day,
                Nex_DMI_kgN_anim_y,Nex_IPCC_kgN_anim_y) |>
  # add share concent
  dplyr::left_join(
    tmp_object_milk@practices$herding$general |>
      # filter for dairy cows
      dplyr::filter(FADN_code_letter == "LCOWDAIR") |>
      dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),
                    share_concent) |>
      # convert in percent
      dplyr::mutate(share_concent = share_concent*100),
    by = tmp_object_milk@traceability$id_cols
  )

# resume table
my_Nexcr_sumup <- my_Nexcr |>
  dplyr::summarise(dplyr::across(
    c(DM_t_anim,
      GE_DMI_MJ_anim_day,GE_IPCC_MJ_anim_day,
      Nex_DMI_kgN_anim_y,Nex_IPCC_kgN_anim_y,
      share_concent),
    list(
      mean = ~ round(mean(.x, na.rm = T),2),
      sd = ~ round(sd(.x, na.rm = T),2)
    ),
    .names = "{.col}_{.fn}"
  ))

```

### Rearing parameters

For dairy cows, we estimate different rearing parameters such as the age at first calving, the renewal rate, or the number of heifers to renew dairy cows.

```{r rearing_param}

# Herd rearing parameters
herd_rear_param = f_herd_rearing_param_cattle(tmp_object_milk) |>
  dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),
                t_1st_calve,rt_LCOWDAIR,LCOWDAIR_Qobs,LCOWDAIR_Fin) |>
  dplyr::mutate(
    # first_calving_age
    first_calving_age = t_1st_calve,
    # renewal_rate
    ## we convert first calving cow number into livestock unit using the heifer livestock unit coefficient
    renewal_rate = mean(LCOWDAIR_Fin*0.8 / LCOWDAIR_Qobs * 100,na.rm = T)
  )


```


### Quantity of purchased feed/head

source ?

```{r}

```


## GHGE

```{r GHGE_output_wheat}

(tmp_object_crop@footprints$GHGE$GHGE_crops)

# Retrieve GHGE estimates for crops, filtering to common wheat only
tmp_ghge_wheat <- tmp_object_crop@footprints$GHGE$GHGE_crops |>
  dplyr::filter(FADN_code_letter == "CWHTC")

# Reshape to long format, keeping per-ha and per-t columns, dropping totals
tmp_ghge_wheat_compo_long <- tmp_ghge_wheat |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("per_ha$|per_t$"),      # keep intensity estimates only
    -dplyr::matches("total_ghg")           # drop aggregated totals
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha$|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "N2O")        ~ "Field N2O Emissions",
      str_detect(component, "ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      TRUE                                ~ "Other Inputs & Processes" # pesticides, seeds, etc.
    ),
    product = "wheat",
    # Build composite indicator label encoding unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",  component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on denominator (area vs. product mass)
    unit = dplyr::case_when(
      grepl("per_ha$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",  component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_wheat_compo <- tmp_ghge_wheat_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_crop@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )


```

```{r GHGE_output_milk}

(tmp_object_milk@footprints$GHGE$GHGE_milk)

# Retrieve GHGE estimates for milk, filtering to cow milk output only
tmp_ghge_milk <- tmp_object_milk@footprints$GHGE$GHGE_milk

# Reshape GHGE milk data to long format, keeping only per-ha (pseudofarm) and per-t columns
tmp_ghge_milk_compo_long <- tmp_ghge_milk |>
  dplyr::select(
    dplyr::all_of(tmp_object_milk@traceability$id_cols),
    dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "milk",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_milk_compo <- tmp_ghge_milk_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_milk@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )


```


```{r GHGE_output_meat}

# select meat farms in France
tmp_object_meat = m_prune_farms(my_object_GHGE, TF8 == 1, COUNTRY == "FRA")

(tmp_object_meat@footprints$GHGE$GHGE_meat)

# Retrieve GHGE estimates for meat, filtering to cow meat output only
tmp_ghge_meat <- tmp_object_meat@footprints$GHGE$GHGE_meat

# Reshape GHGE meat data to long format, keeping only per-ha (pseudofarm) and per-t columns
tmp_ghge_meat_compo_long <- tmp_ghge_meat |>
  dplyr::select(
    dplyr::all_of(tmp_object_meat@traceability$id_cols),
    dplyr::matches("per_ha_pseudofarm|per_t$"),   # keep pseudofarm and per-tonne estimates
    -dplyr::matches("total_ghg|ghge_herd|feed_farm") # drop aggregated/herd-level columns
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("per_ha_pseudofarm|per_t$"),
    names_to = "component"
  ) |>
  dplyr::mutate(
    # Map component names to standardised GHGE categories
    Harmonized_Category = case_when(
      str_detect(component, "(?i)enteric")  ~ "Enteric Fermentation",
      str_detect(component, "(?i)mm|MM")    ~ "Manure Management",
      #      str_detect(component, "(?i)feed")     ~ "Feed Production",
      str_detect(component, "(?i)N2O")        ~ "Field N2O Emissions",
      str_detect(component, "(?i)ferti_prod") ~ "Fertilizer Production",
      str_detect(component, "(?i)diesel|fuel|elec")~ "Energy Use (Diesel/Elec)",
      # TODO: energy use for livestock buildings is not yet accounted for
      TRUE ~ "Other Inputs & Processes"  # pesticides, seeds, transport, post-harvest
    ),
    product = "meat",
    # Build a composite indicator label encoding both unit type and GHGE category
    indicator = dplyr::case_when(
      grepl("per_ha_pseudofarm$", component) ~ paste("GHGE_kgCO2e_ha", Harmonized_Category, sep = "__"),
      grepl("per_t$",            component) ~ paste("GHGE_kgCO2e_t",  Harmonized_Category, sep = "__")
    ),
    # Assign unit label based on whether the estimate is per ha or per tonne of product
    unit = dplyr::case_when(
      grepl("per_ha_pseudofarm$", component) ~ "kgCO2e_per_ha",
      grepl("per_t$",            component) ~ "kgCO2e_per_t"
    )
  )

# Aggregate: sum GHGE components within each farm × category × unit combination
tmp_ghge_meat_compo <- tmp_ghge_meat_compo_long |>
  dplyr::summarise(
    value = sum(value, na.rm = TRUE),
    .by = c(tmp_object_meat@traceability$id_cols, "product", "Harmonized_Category", "indicator", "unit")
  )


```



```{r GHGE_output_farms}

(my_object_GHGE@footprints$GHGE$GHGE_farm)

# Crop farms
tmp_ghge_crop_farm <- tmp_object_crop@footprints$GHGE$GHGE_farm |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("^pseudofarm_total")   # keep pseudofarm estimates
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("pseudofarm_total"),
    names_to = "component"
  ) |>
  # add total area
  dplyr::left_join(
    tmp_object_crop@farm |>
      dplyr::select(dplyr::all_of(tmp_object_crop@traceability$id_cols),SE025),
    by = tmp_object_crop@traceability$id_cols
  ) |>
  # divide by total area
  dplyr::mutate(
    value_per_ha = value / SE025
  ) |>
  # estimate averages
  dplyr::summarise(
    estimate = mean(value_per_ha, na.rm = T),
    .by = 'component'
  ) |>
  dplyr::mutate(
    indicator = paste0(component,"_per_ha"),
    unit = "kgCO2e_per_ha", 
    product = "Field crop farms"
  ) |>
  dplyr::select(
    indicator, unit, product, estimate
  )

# Dairy farms
tmp_ghge_dairy_farm <- tmp_object_milk@footprints$GHGE$GHGE_farm |>
  dplyr::select(
    dplyr::all_of(tmp_object_crop@traceability$id_cols),
    dplyr::matches("^pseudofarm_total")   # keep pseudofarm estimates
  ) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("pseudofarm_total"),
    names_to = "component"
  ) |>
    # add total LU
  dplyr::left_join(
    tmp_object_milk@farm |>
      dplyr::select(dplyr::all_of(tmp_object_milk@traceability$id_cols),SE080),
    by = tmp_object_milk@traceability$id_cols
  ) |>
  # divide by total LU
  dplyr::mutate(
    value_per_LU = value / SE080
  ) |>
  # estimate averages
  dplyr::summarise(
    estimate = mean(value_per_LU, na.rm = T),
    .by = 'component'
  ) |>
  dplyr::mutate(
    indicator = paste0(component,"_per_LU"),
    unit = "kgCO2e_per_LU", 
    product = "Dairy farms"
  ) |>
  dplyr::select(
    indicator, unit, product, estimate
  )


```

## Combine F2F outputs

```{r estimate_table}

# Build agricultural practice indicators for dairy cows from herd nutrition & rearing parameters
F2F_output_agri_pract <- tibble::tribble(
  ~indicator,          ~product,       ~estimate,
  "DMI",               "dairy_cows",   my_Nexcr_sumup$DM_t_anim_mean[[1]],           # dry matter intake (t/animal)
  "GE",                "dairy_cows",   my_Nexcr_sumup$GE_DMI_MJ_anim_day_mean[[1]],  # gross energy intake (MJ/animal/day)
  "Nex",               "dairy_cows",   my_Nexcr_sumup$Nex_DMI_kgN_anim_y_mean[[1]],  # nitrogen excretion (kgN/animal/year)
  "share_concent",     "dairy_cows",   my_Nexcr_sumup$share_concent_mean[[1]],        # concentrate share in diet
  "first_calving_age", "dairy_cows",   mean(herd_rear_param$first_calving_age, na.rm = TRUE) * 12, # convert years to months
  "renewal_rate",      "dairy_cows",   mean(herd_rear_param$renewal_rate, na.rm = TRUE)
) |>
  # Attach units from the literature reference table
  dplyr::left_join(
    lit_agri_pract |> dplyr::select(indicator, unit),
    by = "indicator"
  ) |>
  # add N ferti
  dplyr::bind_rows(
    my_ferti
  )


# Build total GHGE summary (per ha and per tonne) for wheat and milk
F2F_output_ghge <- tibble::tribble(
  ~indicator,        ~unit,             ~product,  ~estimate,
  "GHGE_kgCO2e_ha",  "kgCO2e_per_ha",  "wheat",   mean(tmp_ghge_wheat$total_ghg_crop_kgCO2e_per_ha,                    na.rm = TRUE),
  "GHGE_kgCO2e_t",   "kgCO2e_per_t",   "wheat",   mean(tmp_ghge_wheat$total_ghg_crop_kgCO2e_per_t,                     na.rm = TRUE),
  "GHGE_kgCO2e_ha",  "kgCO2e_per_ha",  "milk",    mean(tmp_ghge_milk$pseudofarm_ghge_herd_kgCO2e_per_ha_pseudofarm,    na.rm = TRUE),
  "GHGE_kgCO2e_t",   "kgCO2e_per_t",   "milk",    mean(tmp_ghge_milk$pseudofarm_ghge_herd_kgCO2e_per_t,                na.rm = TRUE)
) |>
  dplyr::bind_rows(tmp_ghge_crop_farm) |>
  dplyr::bind_rows(tmp_ghge_dairy_farm)

# Compute mean GHGE by harmonized category (component breakdown) for wheat and milk
F2F_output_ghge_compo <- dplyr::bind_rows(
  
  # Wheat: average each GHGE component across farms
  tmp_ghge_wheat_compo |>
    dplyr::summarise(
      estimate = mean(value, na.rm = TRUE),
      .by = c(product, indicator, unit)
    ),
  
  # Milk: average each GHGE component across farms
  tmp_ghge_milk_compo |>
    dplyr::summarise(
      estimate = mean(value, na.rm = TRUE),
      .by = c(product, indicator, unit)
    )
)

# Stack all output tables (practices, total GHGE, GHGE components) into a single flat table
F2F_output <- Reduce(
  function(x, y) dplyr::bind_rows(x, y),
  list(F2F_output_agri_pract,
       F2F_output_ghge,
       F2F_output_ghge_compo)
)

# Render the final output table sorted by product and indicator
F2F_output |> arrange(product, indicator) |> gt()

```

# Validation

We considered a check point validated when the estimates is more than half and less than twice the literature midpoint.

```{r validate}

comparison <- validate_against_literature(
  estimates = F2F_output,
  references = lit_table)

# TODO: check missing share concent
# TODO: check lit value for Bellassen

```

```{r summary}

comparison |>
  count(status) |>
  gt() |>
  tab_header(title = "Validation summary")

```

```{r full_table}

comp_table = comparison |>
  dplyr::filter(!is.na(estimate)) |>
  dplyr::select(indicator, product, unit,
                lit_midpoint, lit_min, lit_max, lit_source,
                estimate, rel_deviation_pct, status) |>
  arrange(product,indicator) |>
  gt() |>
  tab_header(
    title    = "Detailed validation results",
    subtitle = paste("Estimation on France data: ",
                     paste(sort(unique(my_object_GHGE@farm$YEAR)),
                           collapse = ", "))
  ) |>
  fmt_number(columns = c(lit_midpoint, lit_min, lit_max, estimate), decimals = 1) |>
  fmt_number(columns = c(rel_deviation_pct), decimals = 1) |>
  tab_style(
    style     = cell_fill(color = "#fdd"),
    locations = cells_body(rows = status == "\U274C out of range")
  ) |>
  tab_style(
    style     = cell_fill(color = "#dfd"),
    locations = cells_body(rows = status == "\U2705 in range")
  ) |>
  tab_style(
    style     = cell_fill(color = "#ffd"),
    locations = cells_body(rows = grepl("tolerance", status))
  )

comp_table

gtsave(comp_table, filename = "F2F_comp_lit_summary.html")

```

```{r fig-deviation}
#| fig-cap: "Relative deviation from literature midpoint by indicator"
library(ggplot2)

comparison |>
  filter(!is.na(estimate)) |>
  ggplot(aes(x = reorder(paste(indicator, product, sep = " — "),
                         rel_deviation_pct),
             y = rel_deviation_pct,
             fill = status)) +
  geom_col() +
  geom_hline(aes(yintercept = 50), linetype = "dashed", colour = "red") +
  scale_fill_discrete(direction = -1) +
  coord_flip() +
  labs(x = NULL, y = "Relative deviation from midpoint (%)") +
  theme_minimal()

```


