---
title: "FADN2Footprint — Validation Report"
subtitle: "Comparison of package estimates of the pseudoherd with observed herd"
author: "Sarah HUET & Valentin BELLASSEN"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true        
  pdf:
    toc: true
execute:
  echo: true
  warning: false
editor: 
  markdown: 
    wrap: 150
bibliography: ../../vignettes/references.bib
link-citations: true
---

# Purpose

This report compares the outputs of `FADN2Footprint` (v`r packageVersion("FADN2Footprint")`)
against the national inventories.
An estimate is flagged **in range** if it falls within the literature min–max,
**within tolerance** if the relative deviation from the literature midpoint is
below a per-indicator threshold, or **out of range** otherwise.

# Setup

```{r setup}

library(dplyr)
library(tidyr)
library(stringr)
library(gt)

library(readxl)
library(zip)

library(ggplot2)

devtools::load_all()
#library(FADN2Footprint)

```

# Run package

```{r load_data, echo=FALSE, message=TRUE, warning=FALSE}

## Load raw data and run model
#source("validation_FADN_data.R")
source("validation_FADN_data_offline.R")
## 
#load("../../data_raw/FADN_FRA_2018_2021_obj_practices.RData")
#load("../../data_raw/FADN_FRA_2018_2021_obj_GHGE.RData")
#my_object_GHGE <- m_prune_farms(my_object_GHGE,
#                                YEAR %in% c("2018", "2019", "2020"))

object = my_object_GHGE
id_cols = my_object_GHGE@traceability$id_cols

cat("FADN dataset loaded:", nrow(my_object_GHGE@farm), "farm-year observations\n")
cat("Countries covered:", paste(unique(my_object_GHGE@farm$COUNTRY), collapse = ", "), "\n")
cat("Years covered:", paste(sort(unique(my_object_GHGE@farm$YEAR)), collapse = ", "), "\n")

```

# Animal numbers

```{r sum_Q}

pseudoherd_animals <- f_pseudoherd_animals(object)

plot(pseudoherd_animals$Qeq, pseudoherd_animals$Qobs + pseudoherd_animals$Qofffarm)

```


```{r cat_shares}

# herd
herd_total = object@herd |>
  # add SYS02
  dplyr::left_join(
    object@farm |>
      dplyr::select(dplyr::all_of(id_cols), SYS02),
    by = id_cols
  ) |>
  dplyr::summarise(
    category_count = sum(Qobs * SYS02, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR, NUTS2)
  ) |>
  dplyr::mutate(
    category_proportion_national = category_count / sum(category_count),
    .by = c(species, YEAR, NUTS2)
  )  |>
  dplyr::summarise(
    category_proportion_national = mean(category_proportion_national, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  )

# pseudoherd

pseudoherd_animals <- f_pseudoherd_animals(object)

# Total number of animals
pseudoherd_total = pseudoherd_animals |>
  # add SYS02
  dplyr::left_join(
    object@farm |>
      dplyr::select(dplyr::all_of(id_cols), SYS02),
    by = id_cols
  ) |>
  dplyr::summarise(
    category_count = sum(Qeq * SYS02, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  ) |>
  dplyr::mutate(
    category_proportion_pseudoherd = category_count / sum(category_count),
    .by = c(species, YEAR)
  )



# Total number of animals
pseudoherd_total = pseudoherd_animals |>
  dplyr::summarise(
    category_count = sum(Qeq, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  ) |>
  dplyr::mutate(
    category_proportion_pseudoherd = category_count / sum(category_count),
    .by = c(species, YEAR)
  )

# comp

comp <- dplyr::left_join(
  pseudoherd_total,
  herd_total,
  by = c('YEAR', 'FADN_code_letter', 'species')
)

comp |> 
  dplyr::select(YEAR, FADN_code_letter, species, category_proportion_national, category_proportion_pseudoherd) |>
  tidyr::pivot_longer(
    cols = c(category_proportion_national, category_proportion_pseudoherd)
  ) |>
  ggplot() +
  aes(x = FADN_code_letter, y = value, fill = name) +
  geom_col(position = "dodge") +
  #facet_wrap(species) +
  coord_flip()


```

```{r cat_totals_weighted}


# herd
herd_total = object@herd |>
  # add SYS02
  dplyr::left_join(
    object@farm |>
      dplyr::select(dplyr::all_of(id_cols), SYS02),
    by = id_cols
  ) |>
  dplyr::summarise(
    category_count_national = sum(Qobs * SYS02, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  ) 


# pseudoherd

pseudoherd_animals <- f_pseudoherd_animals(object)

# Total number of animals - weighted
pseudoherd_total = pseudoherd_animals |>
  # add SYS02
  dplyr::left_join(
    object@farm |>
      dplyr::select(dplyr::all_of(id_cols), SYS02),
    by = id_cols
  ) |>
  dplyr::summarise(
    category_count_pseudoherd = sum(Qeq * SYS02, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  )

# Total number of animals - unweighted
#pseudoherd_total = pseudoherd_animals |>
#  dplyr::summarise(
#    category_count_pseudoherd = sum(Qeq, na.rm = T),
#    .by = c(FADN_code_letter, species, YEAR)
#  )


# remove export
Agreste_Echanges_extérieurs_d_animaux_de_boucherie_vivants <- read_csv("Agreste_Echanges extérieurs d’animaux de boucherie vivants.csv")

exports <- Agreste_Echanges_extérieurs_d_animaux_de_boucherie_vivants |>
  tidyr::pivot_longer(cols = -c("YEAR","Indicateur"),
                      names_to = "agreste_cat",
                      values_to = "value") |>
  tidyr::pivot_wider(
    id_cols = c(YEAR, agreste_cat),
    names_from = Indicateur
  ) |>
  dplyr::mutate(
    YEAR = as.character(YEAR),
    #balance = `Exportations tous animaux (tête)`
    balance = `Exportations tous animaux (tête)` - `Importations tous animaux (tête)`
  ) |>
  # match category
  dplyr::mutate(
    FADN_code_letter = dplyr::case_when(
      agreste_cat == "Veaux de boucherie" ~ "LBOV1",
      agreste_cat == "Broutards légers" ~ "LBOV1_2M",
      agreste_cat == "Génisses" ~ "LHEIFFAT",
      agreste_cat == "Vaches" ~ "LCOWOTH",
      agreste_cat == "Gros bovins mâles" ~ " LBOV2"
    )
  )

pseudoherd_balanced <- pseudoherd_total |>
  # add export balance
  dplyr::left_join(exports |>
                     dplyr::select(FADN_code_letter, YEAR, balance),
                   by = c('FADN_code_letter', 'YEAR')) |>
  # balance
  dplyr::mutate(
    category_count_pseudoherd_balance = category_count_pseudoherd - dplyr::coalesce(balance, 0)
  )

# comp
dplyr::left_join(
  pseudoherd_balanced,
  herd_total,
  by = c('YEAR', 'FADN_code_letter', 'species')
) |> 
  #dplyr::filter(species == "cattle") |>
  dplyr::select(YEAR, FADN_code_letter, species, dplyr::matches("count")) |>
  tidyr::pivot_longer(
    cols = c(category_count_national, category_count_pseudoherd, category_count_pseudoherd_balance)
    #cols = c(category_count_national, category_count_pseudoherd)
  ) |>
  ggplot() +
  aes(x = name, y = value, fill = FADN_code_letter) +
  geom_col(position = "fill") +
  facet_grid(vars(species)) +
  coord_flip() +
  scale_fill_viridis_d() 

```

https://agreste.agriculture.gouv.fr/agreste-saiku/?plugin=true&query=query/open/COMEXABVIV#query/open/COMEXABVIV

```{r share_stage}

# Qobs_meat

cat_juveniles <- c("LBOV1")
cat_fattening <- c("LBOV1_2M", "LBOV2", "LHEIFFAT", "LBOV1_2F_fattening")
cat_breeders  <- c("LHEIFBRE", "LCOWOTH", "LBOV1_2F_breeders")

herd_activities <- f_herd_activities(object)

# estimate shares
share_Qobs_meat <- herd_activities |>
  # add NUTS2 and SYS02
  dplyr::left_join(object@farm |>
                     dplyr::select(dplyr::all_of(id_cols), NUTS2, SYS02),
                   by = id_cols) |>
  # sum all animals per category in each rearing stage at the NUTS2 level
  dplyr::summarise(
    Qobs_meat_NUTS2_cat = sum(Qobs_meat, na.rm = T),
    .by = c(COUNTRY, NUTS2, FADN_code_letter)
  ) |>
  # sum all animals per category in each rearing stage at the country level
  dplyr::mutate(
    Qobs_meat_COUNTRY_cat = sum(Qobs_meat_NUTS2_cat, na.rm = T),
    .by = c(COUNTRY, FADN_code_letter)
  ) |>
  # add rearing stage
  dplyr::mutate(
    stage = dplyr::case_when(
      FADN_code_letter %in% cat_juveniles ~ "j",
      FADN_code_letter %in% cat_fattening  ~ "f",
      FADN_code_letter %in% cat_breeders   ~ "b",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::filter(!is.na(stage)) |>
  # sum all animals at the NUTS2 level (per stage)
  dplyr::mutate(
    Qobs_meat_NUTS2_stage = sum(Qobs_meat_NUTS2_cat, na.rm = T),
    .by = c(NUTS2, stage)
  ) |>
  # sum all animals at the country level (per stage)
  dplyr::mutate(
    Qobs_meat_COUNTRY_stage = sum(Qobs_meat_NUTS2_cat, na.rm = T),
    .by = c(COUNTRY, stage)
  ) |>
  # estimate share of animal per category at the NUTS2 level
  dplyr::mutate(
    share_NUTS2_Qobs_meat = Qobs_meat_NUTS2_cat / Qobs_meat_NUTS2_stage,
    share_COUNTRY_Qobs_meat = Qobs_meat_COUNTRY_cat / Qobs_meat_COUNTRY_stage
  )


# Qobs
# estimate shares
share_Qobs <- herd_activities |>
  # add NUTS2 and SYS02
  dplyr::left_join(object@farm |>
                     dplyr::select(dplyr::all_of(id_cols), NUTS2, SYS02),
                   by = id_cols) |>
  # sum all animals per category in each rearing stage at the NUTS2 level
  dplyr::summarise(
    Qobs_NUTS2_cat = sum(Qobs, na.rm = T),
    .by = c(COUNTRY, NUTS2, FADN_code_letter)
  ) |>
  # sum all animals per category in each rearing stage at the country level
  dplyr::mutate(
    Qobs_COUNTRY_cat = sum(Qobs_NUTS2_cat, na.rm = T),
    .by = c(COUNTRY, FADN_code_letter)
  ) |>
  # add rearing stage
  dplyr::mutate(
    stage = dplyr::case_when(
      FADN_code_letter %in% cat_juveniles ~ "j",
      FADN_code_letter %in% cat_fattening  ~ "f",
      FADN_code_letter %in% cat_breeders   ~ "b",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::filter(!is.na(stage)) |>
  # sum all animals at the NUTS2 level (per stage)
  dplyr::mutate(
    Qobs_NUTS2_stage = sum(Qobs_NUTS2_cat, na.rm = T),
    .by = c(NUTS2, stage)
  ) |>
  # sum all animals at the country level (per stage)
  dplyr::mutate(
    Qobs_COUNTRY_stage = sum(Qobs_NUTS2_cat, na.rm = T),
    .by = c(COUNTRY, stage)
  ) |>
  # estimate share of animal per category at the NUTS2 level
  dplyr::mutate(
    share_NUTS2_Qobs = Qobs_NUTS2_cat / Qobs_NUTS2_stage,
    share_COUNTRY_Qobs = Qobs_COUNTRY_cat / Qobs_COUNTRY_stage
  )

# comp

comp <- dplyr::left_join(
  share_Qobs,
  share_Qobs_meat,
  by = c('COUNTRY', 'NUTS2', 'FADN_code_letter', 'stage')
)

comp |> 
  dplyr::select(FADN_code_letter, dplyr::matches("share")) |>
  tidyr::pivot_longer(
    cols = dplyr::matches("share")
  ) |>
  ggplot() +
  aes(x = name, y = value, fill = FADN_code_letter) +
  geom_col(position = "fill") +
  #facet_grid(vars(species)) +
  coord_flip() +
  scale_fill_viridis_d() 



```


# Emissions

```{r sum_ghge}

pseudoherd_ghge = object@footprints$GHGE$GHGE_pseudoherd

plot(pseudoherd_ghge$farm_ghge_herd_kgCO2e_livcat_pseudoherd, pseudoherd_ghge$farm_ghge_herd_kgCO2e_livcat + pseudoherd_ghge$farm_ghge_herd_kgCO2e_livcat_offfarm)

```


```{r ghge_per_anim_offfarm}

herd_ghge = object@footprints$GHGE$GHGE_herd |>
  dplyr::filter(Qobs >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_herd = mean(farm_ghge_herd_kgCO2e_per_anim, na.rm = TRUE),
    .by = FADN_code_letter
  )

pseudoherd_ghge = object@footprints$GHGE$GHGE_pseudoherd |>
  dplyr::filter(Qeq >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_offfarm = mean(farm_ghge_herd_kgCO2e_livcat_offfarm / Qofffarm, na.rm = TRUE),
    .by = FADN_code_letter
  )

dplyr::left_join(
  herd_ghge,
  pseudoherd_ghge,
  by = c('FADN_code_letter')
) |> 
  dplyr::select(FADN_code_letter, farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_offfarm) |>
  tidyr::pivot_longer(
    cols = c(farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_offfarm)
  ) |>
  ggplot() +
  aes(x = FADN_code_letter, y = value, fill = name) +
  geom_col(position = "dodge") +
  coord_flip()


```

```{r ghge_per_anim_Qeq}

herd_ghge = object@footprints$GHGE$GHGE_herd |>
  dplyr::filter(Qobs >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_herd = mean(farm_ghge_herd_kgCO2e_per_anim, na.rm = TRUE),
    .by = FADN_code_letter
  )

pseudoherd_ghge = object@footprints$GHGE$GHGE_pseudoherd |>
  dplyr::filter(Qeq >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_pseudoherd = mean(farm_ghge_herd_kgCO2e_livcat_pseudoherd / Qeq, na.rm = TRUE),
    .by = FADN_code_letter
  )

dplyr::left_join(
  herd_ghge,
  pseudoherd_ghge,
  by = c('FADN_code_letter')
) |> 
  dplyr::select(FADN_code_letter, farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_pseudoherd) |>
  tidyr::pivot_longer(
    cols = c(farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_pseudoherd)
  ) |>
  ggplot() +
  aes(x = FADN_code_letter, y = value, fill = name) +
  geom_col(position = "dodge") +
  coord_flip()


```

```{r ghge_milk}

object@footprints$GHGE$GHGE_milk |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_t = mean(farm_ghge_herd_kgCO2e_per_t, na.rm= TRUE)
  )

object@footprints$GHGE$GHGE_pseudoherd_milk |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_pseudoherd_per_t = mean(farm_ghge_herd_kgCO2e_pseudoherd_per_t, na.rm= TRUE)
  )


```

```{r ghge_meat}

object@footprints$GHGE$GHGE_meat |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_t = mean(farm_ghge_herd_kgCO2e_per_t, na.rm= TRUE),
    .by = output
  )

object@footprints$GHGE$GHGE_pseudoherd_meat |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_pseudoherd_per_t = mean(farm_ghge_herd_kgCO2e_pseudoherd_per_t, na.rm= TRUE),
    .by = output
  )


```


