---
title: "FADN2Footprint — Validation Report for Pseudoherd"
subtitle: "Comparison of package estimates of the pseudoherd with observed herd"
author: "Sarah HUET & Valentin BELLASSEN"
date: today
editor: 
  markdown: 
    wrap: 150
---

# Setup

```{r setup}

library(dplyr)
library(tidyr)
library(stringr)
library(gt)

library(readxl)
library(zip)

library(ggplot2)

devtools::load_all()
#library(FADN2Footprint)

wd_path = "C:/Users/srhuet/OneDrive/Research/GitHub/FADN2Footprint"

```
# Purpose

This report compares the outputs of `FADN2Footprint` (v`r packageVersion("FADN2Footprint")`)
against the national inventories.
An estimate is flagged **in range** if it falls within the literature min–max,
**within tolerance** if the relative deviation from the literature midpoint is
below a per-indicator threshold, or **out of range** otherwise.

# Run package

```{r load_data, echo=FALSE, message=TRUE, warning=FALSE}

## Load raw data and run model
#source("validation_FADN_data.R")
#source("inst/validation/validation_FADN_data_offline.R")
##
#load("data_raw/FADN_FRA_2018_2021_obj.RData") 
#load("../../data_raw/FADN_FRA_2018_2021_obj_practices.RData")
load(paste0(wd_path, "/data_raw/FADN_FRA_2018_2021_obj_GHGE.RData"))
#my_object_GHGE <- m_prune_farms(my_object_GHGE,
#                                YEAR %in% c("2018", "2019", "2020"))

object = my_object_GHGE
id_cols = my_object_GHGE@traceability$id_cols

cat("FADN dataset loaded:", nrow(my_object_GHGE@farm), "farm-year observations\n")
cat("Countries covered:", paste(unique(my_object_GHGE@farm$COUNTRY), collapse = ", "), "\n")
cat("Years covered:", paste(sort(unique(my_object_GHGE@farm$YEAR)), collapse = ", "), "\n")

```

# Animal numbers

## Number of calves sold for slaughter

https://www.franceagrimer.fr/sites/default/files/2026-02/28_FICHE_FILIERE_VEAU_2026.pdf
entre 1200 et 1300 miliers de tête de veaux par an en France entre 2018 et 2020

```{r calves_slaughter}

object@herd |>
  dplyr::filter(FADN_code_letter == "LBOV1") |>
  # add sys02
  dplyr::left_join(object@farm |>
                     dplyr::select(dplyr::all_of(id_cols), SYS02),
                   by = id_cols) |>
  # sum
  dplyr::summarise(sum_calves_Qobs_SYS02 = sum(Qobs * SYS02, na.rm = T),
                   sum_calves_Qobs = sum(Qobs, na.rm = T),
                   sum_calves_SSN_SYS02 = sum(SSN * SYS02, na.rm = T),
                   sum_calves_SSN = sum(SSN, na.rm = T),
                   .by = YEAR)

```

## Sum per livestock category

```{r sum_per_cat}

object@herd |>
  # add sys02
  dplyr::left_join(object@farm |>
                     dplyr::select(dplyr::all_of(id_cols), SYS02),
                   by = id_cols) |>
  # sum
  dplyr::summarise(sum_Qobs_SYS02 = sum(Qobs * SYS02, na.rm = T),
                   sum_Qobs = sum(Qobs, na.rm = T),
                   sum_SSN_SYS02 = sum(SSN * SYS02, na.rm = T),
                   sum_SSN = sum(SSN, na.rm = T),
                   .by = c(YEAR, FADN_code_letter)) |>
  view()


```



## Share between categories

To ensure that the pseudoherd is consistent with the observed herd,  we compare the share of each livestock category in the total number of animals in the observed herd and in the pseudoherd.
We add to the observed herd the number of animals exported, as the pseudoherd is meant to account for all animals involved in the production, including those exported.


```{r cat_add_export}

# herd
herd_total = object@herd |>
  # add SYS02
  dplyr::left_join(
    object@farm |>
      dplyr::select(dplyr::all_of(id_cols), SYS02),
    by = id_cols
  ) |>
  dplyr::summarise(
    category_count_national = sum(Qobs * SYS02, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  ) 


# pseudoherd

pseudoherd_animals <- f_pseudoherd_animals(object)

# Total number of animals - weighted
pseudoherd_total = pseudoherd_animals |>
  # add SYS02
  dplyr::left_join(
    object@farm |>
      dplyr::select(dplyr::all_of(id_cols), SYS02),
    by = id_cols
  ) |>
  dplyr::summarise(
    category_count_pseudoherd = sum(Qeq * SYS02, na.rm = T),
    .by = c(FADN_code_letter, species, YEAR)
  )

# Total number of animals - unweighted
#pseudoherd_total = pseudoherd_animals |>
#  dplyr::summarise(
#    category_count_pseudoherd = sum(Qeq, na.rm = T),
#    .by = c(FADN_code_letter, species, YEAR)
#  )

# add export
library(readr)
Agreste_Echanges_extérieurs_d_animaux_de_boucherie_vivants <- read_csv(paste0(wd_path, "/inst/validation/Agreste_Echanges extérieurs d’animaux de boucherie vivants.csv"))

exports <- Agreste_Echanges_extérieurs_d_animaux_de_boucherie_vivants |>
  tidyr::pivot_longer(cols = -c("YEAR","Indicateur"),
                      names_to = "agreste_cat",
                      values_to = "value") |>
  tidyr::pivot_wider(
    id_cols = c(YEAR, agreste_cat),
    names_from = Indicateur
  ) |>
  dplyr::mutate(
    YEAR = as.character(YEAR),
    balance = `Exportations tous animaux (tête)`
    #balance = `Exportations tous animaux (tête)` - `Importations tous animaux (tête)`
  ) |>
  # match category
  dplyr::mutate(
    FADN_code_letter = dplyr::case_when(
      agreste_cat == "Veaux de boucherie" ~ "LBOV1",
      agreste_cat == "Broutards légers" ~ "LBOV1_2M",
      agreste_cat == "Génisses" ~ "LHEIFFAT",
      agreste_cat == "Vaches" ~ "LCOWOTH",
      agreste_cat == "Gros bovins mâles" ~ "LBOV2"
    )
  )

herd_balanced <- herd_total |>
  # add export balance
  dplyr::left_join(exports |>
                     dplyr::select(FADN_code_letter, YEAR, balance),
                   by = c('FADN_code_letter', 'YEAR')) |>
  # balance
  dplyr::mutate(
    category_count_national_balance = category_count_national + dplyr::coalesce(balance, 0)
  )

# comp
dplyr::left_join(
  pseudoherd_total,
  herd_balanced,
  by = c('YEAR', 'FADN_code_letter', 'species')
) |> 
  #dplyr::filter(species == "cattle") |>
  dplyr::select(YEAR, FADN_code_letter, species, dplyr::matches("count")) |>
  tidyr::pivot_longer(
    #cols = c(category_count_national, category_count_pseudoherd, category_count_pseudoherd_balance)
    cols = c(category_count_national_balance, category_count_pseudoherd)
  ) |>
  ggplot() +
  aes(x = name, y = value, fill = FADN_code_letter) +
  geom_col(position = "fill") +
  facet_grid(vars(species)) +
  coord_flip() +
  paletteer::scale_fill_paletteer_d("ggthemes::Tableau_20")

```

# Emissions

We compare the emissions per animal in the observed herd and in the pseudoherd, for each livestock category.
First, we look at the emission of the off-farm animals.

```{r ghge_per_anim_offfarm}

herd_ghge = object@footprints$GHGE$GHGE_herd |>
  dplyr::filter(Qobs >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_herd = mean(farm_ghge_herd_kgCO2e_per_anim, na.rm = TRUE),
    .by = c(FADN_code_letter, YEAR)
  )

pseudoherd_ghge = object@footprints$GHGE$GHGE_pseudoherd |>
  dplyr::filter(Qeq >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_offfarm = mean(farm_ghge_herd_kgCO2e_livcat_offfarm / Qofffarm, na.rm = TRUE),
    .by = c(FADN_code_letter, YEAR)
  )

dplyr::left_join(
  herd_ghge,
  pseudoherd_ghge,
  by = c('FADN_code_letter', 'YEAR')
) |> 
  dplyr::select(FADN_code_letter, farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_offfarm) |>
  tidyr::pivot_longer(
    cols = c(farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_offfarm)
  ) |>
  ggplot() +
  aes(x = FADN_code_letter, y = value, fill = name) +
  geom_col(position = "dodge") +
  coord_flip()


```
Then, we look at the emission of all animals in the pseudoherd, including those on-farm and off-farm.

```{r ghge_per_anim_Qeq}

herd_ghge = object@footprints$GHGE$GHGE_herd |>
  dplyr::filter(Qobs >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_herd = mean(farm_ghge_herd_kgCO2e_per_anim, na.rm = TRUE),
    .by = c(FADN_code_letter, YEAR)
  )

pseudoherd_ghge = object@footprints$GHGE$GHGE_pseudoherd |>
  dplyr::filter(Qeq >= 1) |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_anim_pseudoherd = mean(farm_ghge_herd_kgCO2e_livcat_pseudoherd / Qeq, na.rm = TRUE),
    .by = c(FADN_code_letter, YEAR)
  )

dplyr::left_join(
  herd_ghge,
  pseudoherd_ghge,
  by = c('FADN_code_letter', 'YEAR')
) |> 
  dplyr::select(FADN_code_letter, farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_pseudoherd) |>
  tidyr::pivot_longer(
    cols = c(farm_ghge_herd_kgCO2e_per_anim_herd, farm_ghge_herd_kgCO2e_per_anim_pseudoherd)
  ) |>
  ggplot() +
  aes(x = FADN_code_letter, y = value, fill = name) +
  geom_col(position = "dodge") +
  coord_flip()


```

```{r ghge_milk}

object@footprints$GHGE$GHGE_milk |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_t = mean(farm_ghge_herd_kgCO2e_per_t, na.rm= TRUE)
  )

object@footprints$GHGE$GHGE_pseudoherd_milk |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_pseudoherd_per_t = mean(farm_ghge_herd_kgCO2e_pseudoherd_per_t, na.rm= TRUE)
  )


```

```{r ghge_meat}

object@footprints$GHGE$GHGE_meat |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_per_t = mean(farm_ghge_herd_kgCO2e_per_t, na.rm= TRUE),
    .by = c(output, YEAR)
  ) |>
  dplyr::arrange(output, YEAR)

object@footprints$GHGE$GHGE_pseudoherd_meat |>
  dplyr::summarise(
    farm_ghge_herd_kgCO2e_pseudoherd_per_t = mean(farm_ghge_herd_kgCO2e_pseudoherd_per_t, na.rm= TRUE),
    .by = c(output, YEAR)
  ) |>
  dplyr::arrange(output, YEAR)


```


